---
title: "The effect of donation activity dwarfs the effect of lifestyle, diet and iron supplementation on blood donor iron stores - Analysis code"
author: "Muriel Lobier"
output: github_document
bibliography: index.bib
---

# Executive summary 

This document includes all code necessary to run the analysis and produce the figures of : *The effect of donation activity dwarfs the effect of lifestyle, diet and iron supplementation on donor iron status.* Lobier, Nittymäki, Castren, Palokongas, Partanen and Arvas. 

The analysis has  three parts. In a first part we describe the study population and compute the prevalence of iron deficiency and anemia. In a second part, we use multivariate linear regression to identify the factors that co-vary with iron levels (using ferritin and sTfR) in the blood donor population. In a third part, we test whether ferritin levels affect donor self-reported health using ordinal logistic regression. 


```{r setup, , message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
# library(janitor)
library(tableone)
library(stargazer)
library(gridExtra)
library(cowplot)
library(broom)
library(GGally)
library(ggfortify)
# library(sjPlot)
library(knitr)
library(lubridate)
library(HDInterval)
library(car)
library(relaimpo)
library(sfsmisc)
library(MASS)
# library(kableExtra)
library(brms)
library(rstan)
library(rms)
library(modelr)
library(tidybayes)
library(ordinal)
library(sjmisc)
library(sjlabelled)
# library(lazerhawk)
```


# Data loading and preparation

```{r , warning=FALSE}

load("../data/r02.fd.bd.all.rdata")

indiv_donations_data <- output

rm(output)
```


## Load Data


There are `r indiv_donations_data %>% distinct(donor) %>% nrow()` donors enrolled in the study. 

There are `r indiv_donations_data %>%  filter(!is.na(Ferritin) & !is.na(Hb_v) ) %>% distinct(donor) %>% nrow()` left once we remove donors that have no Ferritin or no Hb measurement for any donation event. 



```{r , warning=FALSE}
# Get sex for each donor
blood_data_summary <- indiv_donations_data %>% 
  group_by(donor) %>% 
  summarise(sex=first(gender)) 

# Get values for first study donation with the required measurements 
# donation (regardless of donation type)
# We remove events with no Ferritin and hb_v to be inline with from when 
# the nb of previous donations were counted.


blood_values_init <- indiv_donations_data %>% 
  filter(!is.na(Ferritin) & !is.na(Hb_v) ) %>% 
  group_by(donor) %>% 
  filter(date == min(date)) %>% 
  dplyr::select(age,TransferrinR, Ferritin,Hb_v,CRP,DaysToPreviousFB, MCV, donor) %>% 
  ungroup()


blood_data_summary <- blood_data_summary %>% 
  full_join(blood_values_init, by = "donor") 

# We now add the demographic data
load("../data/r02ds.donorData.rdata")
dons <- output
rm(output)


blood_data_summary <- dons %>%
  inner_join(blood_data_summary, by = "donor") %>%
  mutate(all_study_FB_donation_count = NonFinnDonorDonationCount_FB + YesFinnDonorDonationCount_FB) 
```


## Demographic group specification and assignment

There are `r blood_data_summary %>% filter(sex == "Women" & is.na(QR79)) %>%  nrow()` women with no  answer to the question regarding their menstrual status. Of these, `r blood_data_summary %>% filter(sex == "Women" & is.na(QR79) & age > 65) %>%  nrow()` are older than 65. We impute these to  no period so they can be included in the post-menopausal women. Other women are all under 55, thus their menstrual status cannot be ascertained. 


```{r}
older_donors <- filter(blood_data_summary, 
                       sex == "Women" & is.na(QR79) & age > 55)$donor

blood_data_summary <- blood_data_summary %>% 
  mutate(QR79 = ifelse(donor %in% older_donors, "no_period", as.character(QR79))) 
```


Female donors with no menstruation response after imputation are removed: `r blood_data_summary %>% filter(sex == "Women" & is.na(QR79)) %>% nrow()` donors.


```{r}
nb_women_no_mesntruation_response <- blood_data_summary %>% filter(sex == "Women" & is.na(QR79)) %>% nrow()


# remove female donors with no menstruation response
blood_data_summary_final <- blood_data_summary %>% 
  mutate(mens_ok = case_when(
    sex == "Men" ~ "Men", 
    sex == "Women" & !is.na(QR79)  ~ "Women",
    TRUE ~ "NA")) %>% 
  filter(mens_ok != "NA") 

# We define the women's groups: 
n_women_removed <- blood_data_summary_final %>% 
  mutate(group = case_when(
    sex == "Men" ~ "Men", 
    sex == "Women" & (QR79 == "no_period" & age > 45) ~ "Post_menopause_women",
    sex == "Women" & (QR79 == "no_period" & age <= 45) ~ "Women_pre_menop_no_mens",
    sex == "Women" & (QR79 == "irregular_period" | QR79 == "regular_period") ~ "Pre_menopause_women",
    TRUE ~ "NA")) %>% 
  group_by(group) %>% 
  filter(group ==  "Women_pre_menop_no_mens") %>% 
  nrow()

```

We now define the following groups:

+ pre-menopausal: regular or irregular menstruation reported
+ post-menopausal: no menstruation reported and age equal to superior to 45
+ donors younger than 45 and no reported menstruation are excluded (`r n_women_removed` donors)



```{r}
# We define the women's groups: 
blood_data_summary_final <- blood_data_summary_final %>% 
  mutate(group = case_when(
    sex == "Men" ~ "Men", 
    sex == "Women" & (QR79 == "no_period" & age > 45) ~ "Post_menopause_women",
    sex == "Women" & (QR79 == "no_period" & age <= 45) ~ "Women_pre_menop_no_mens",
    sex == "Women" & (QR79 == "irregular_period" | QR79 == "regular_period") ~ "Pre_menopause_women",
    TRUE ~ "NA")) %>% 
  filter(group != "Women_pre_menop_no_mens" & group != "NA") %>% 
  droplevels() %>% 
  mutate(group = ordered(group, levels =  c("Pre_menopause_women", "Post_menopause_women", "Men"))) 
```



## Remove donors with missing data

### Blood measures and donor demographics

We remove donors for which we do not have:

* CRP or sTfR: `r blood_data_summary_final %>% filter(is.na(CRP) | is.na(TransferrinR)) %>% nrow()` donors

```{r}
nb_NA_removed <- blood_data_summary_final %>%
  filter(is.na(TransferrinR) | is.na(CRP) ) %>% 
  nrow()

blood_data_summary_final <- blood_data_summary_final %>% 
  drop_na(TransferrinR, CRP) 

```


### Donation history

We remove `r blood_data_summary_final %>% filter(is.na(DaysToPreviousFB)) %>% nrow()` donors who have not donated previously (They are missing the Nb of days since last FB donation variable)

```{r}
new_donors_data <- blood_data_summary_final %>% 
  filter(is.na(DaysToPreviousFB)) %>% 
    mutate(group = case_when(group == "Pre_menopause_women" ~ "Pre-menopausal women",
                   group == "Post_menopause_women" ~ "Post-menopausal women",
                   group == "Men" ~ "Men" ,
                   TRUE~ "NA"),
         group = ordered(group)) 

blood_data_summary_final <- blood_data_summary_final %>% 
  drop_na(DaysToPreviousFB)
```

### BMI

We remove `r blood_data_summary_final %>% filter(is.na(BMI)) %>% nrow()` donors for whom we do not have the BMI data:

```{r}
nb_removed_questionnaire <- nb_women_no_mesntruation_response +
                            blood_data_summary_final %>% filter(is.na(BMI)) %>% nrow() 

blood_data_summary_final <- blood_data_summary_final %>% 
  filter(!is.na(BMI))
```

### Smoking

We remove `r blood_data_summary_final %>% filter(is.na(QR54)) %>% nrow()` donors  who  did not answer the smoking question:

```{r}
nb_removed_questionnaire <- nb_removed_questionnaire +
                            blood_data_summary_final %>% filter(is.na(QR54)) %>% nrow() 

blood_data_summary_final <- blood_data_summary_final %>% 
  filter(!is.na(QR54))
```

### Pregnancy

We remove `r blood_data_summary_final %>% filter((group != "Men" & is.na(QR83)))%>% nrow()` female donors who did not answer the pregnancy question: 

```{r}
nb_removed_questionnaire <- nb_removed_questionnaire +
                            blood_data_summary_final %>% filter((group != "Men" & is.na(QR83))) %>% nrow 

blood_data_summary_final <- blood_data_summary_final %>% 
  filter(!(group != "Men" & is.na(QR83)))
```

### Iron supplementation 

We remove `r blood_data_summary_final %>% mutate(iron_comp_c = ifelse(iron_supp == FALSE, 0, iron_comp_c )) %>%  filter(is.na(iron_comp_c)) %>% nrow()` donors that did not answer properly to both questions if applicable.

For the modelling, we impute a 0 (no supplementation) to iron_comp_c when the donor reports not being offered iron supplementation.

```{r}
nb_removed_questionnaire <- nb_removed_questionnaire +
                            blood_data_summary_final %>% 
                              mutate(iron_comp_c = ifelse(iron_supp == FALSE, 0, iron_comp_c )) %>% 
                              filter(is.na(iron_comp_c)) %>% 
                              nrow() 

blood_data_summary_final <- blood_data_summary_final %>% 
  mutate(iron_comp_c = ifelse(iron_supp == FALSE, 0, iron_comp_c )) %>% 
  filter(!is.na(iron_comp_c))
```

### Food and drink intake

* QR40 how often do you eat meat
* QR44: how often do you eat fruit and berries
* QR46: how often do you drink fruit juice
* QR45: how often do you eat vegetables
* QR48: how often do you drink milk 
* QR49: how often do you drink coffee
* QR50: how often do you drink tea
* QR51: how often do you drink beer
* QR52: how often do you drink wine 
* QR53: how often do you drink strong alcohol 

```{r message=FALSE, warning=FALSE}

donors_to_remove <- blood_data_summary_final %>% 
  dplyr::select(donor, group, QR40,QR44, QR46, QR45, QR48, QR49, QR50,QR51, QR52, QR53) %>% 
  gather(key= question, value = answer, -donor,-group) %>% 
  filter(is.na(answer)) %>% 
  dplyr::select(group, donor) %>% 
  distinct(donor, group) 


blood_data_summary_final <- blood_data_summary_final %>% 
  filter(!donor %in% donors_to_remove$donor) 

nb_removed_questionnaire <- nb_removed_questionnaire +
                            donors_to_remove %>% nrow()


```

We remove `r donors_to_remove %>% nrow()` donors that did not answer these questions. 

### Health self assessment

* QR17 (X5) how does your health feel in general ?

We remove `r blood_data_summary_final %>% filter(is.na(QR17)) %>% nrow()` donors  who  did not answer the health self-assessment question:

```{r}
nb_removed_questionnaire <- nb_removed_questionnaire +
                            blood_data_summary_final %>% filter(is.na(QR17)) %>% nrow() 

blood_data_summary_final <- blood_data_summary_final %>% 
  filter(!is.na(QR17))
```


### Physical activity

We remove `r blood_data_summary_final %>% filter(is.na(QR60) | is.na(QR61)) %>% nrow()` donors  who  did not answer the physical activity questions:


```{r}

nb_removed_questionnaire <- nb_removed_questionnaire +
                            blood_data_summary_final %>% filter(is.na(QR60) | is.na(QR61)) %>% nrow()

blood_data_summary_final <- blood_data_summary_final %>% 
  drop_na(QR60, QR61)

```

### Total number of donors removed for missing questionnaire answers

The total number of donors removed because of missing answers in the questionnaire is `r nb_removed_questionnaire`

## Remove donors with extreme physiological measures

As decided previously, we remove data according the following criteria:

* BMI > 50
* CRP > 30
* Ferritin > 400 

This amounts to `r blood_data_summary_final %>% filter(BMI >= 50 | CRP >= 30 | Ferritin >= 400) %>% nrow()` donors that are removed. 


```{r}
 blood_data_summary_final <- blood_data_summary_final %>% 
   filter(BMI < 50 & CRP < 30 & Ferritin < 400)
```

## Final group N

```{r}
blood_data_summary_final %>% 
   mutate(group = dplyr::recode(group, Pre_menopause_women = "Pre-menopausal women",
         Post_menopause_women = "Post-menopausal women")) %>% 
  group_by(group) %>% 
  summarise ( N = n()) %>% 
  kable() 
 
```


## Recoding of variables

```{r}

# We keep only the variables that are used in the analysis

blood_data_summary_final <- blood_data_summary_final %>% 
  dplyr::select(donor,group, sex, age,TwoYearsFromStartCount_FB,DaysToPreviousFB, Hb_v,
       Ferritin, CRP, TransferrinR, MCV, DaysToPreviousFB,QR40,QR44, QR46, QR45, QR48,
       QR49, QR50,QR51, QR52, QR53, 
       QR17, QR54, QR60, QR61, iron_comp_c, BMI,QR83, iron_supp) %>% 
  rename(red_meat = QR40,
         fruit_berries = QR44,
         vegetables = QR45,
         fruit_juices = QR46,
         milk = QR48,
         coffee = QR49,
         tea = QR50,
         beer = QR51,
         wine = QR52,
         liquor = QR53,
         pregnancy = QR83,
         self_perceived_health = QR17,
         physical_act_daily = QR60,
         physical_act_freq = QR61) %>% 
    mutate(smoking = ifelse(QR54 == "no", "no", "yes"),
         smoking = ordered(smoking, levels = c( "no",  "yes")),
         pregnancy = ordered(pregnancy, levels = c( "no",  "yes")))


```

We recode the following variables:

* Food and non-alcoholic beverages
* Alcoholoc beverages
* Smoking
* Pregnancy

We use a linear scale from 1 to 4 for both food/non-alcoholic drinks and for alcoholic drinks. However, since the proposed answers were different, the mappings between numerical value and answers are different between these variable groups. FOr alcoholic beverages, we used the same scale as used in Rigas 2014. For Non alcoholic beverages and food, we mapped the 6 pssible response the 4 codes. 

* both food/non-alcoholic drinks:
    * "never" ~ 1,
    * "less_than_once_weekly" ~ 2,
    * "1.3_week" or "4.6_week"~ 3,
    * "daily" or "several_daily" ~ 4,

* Alcoholic drinks:
    * "never" or "very_rarely" ~ 1,
    * "a_few_per_month" ~ 2,
    * "a_few_per_week" ~ 3,
    * "daily_or_almost" ~ 4,

* Smoking:
    * "no" ~"no"
    * "sometimes" or "daily"  ~ "yes"
   
* QR83: have you given birth ?

```{r}
# Recode food and non-alcoholic beverages
food <- blood_data_summary_final %>% 
  dplyr::select(donor,red_meat, fruit_berries,vegetables,fruit_juices,milk, coffee, tea ) %>% 
  gather(key = key, value =  value, -donor) %>% 
  mutate(linear_value = case_when(value == "never" ~ 1,
    value == "less_than_once_weekly" ~ 2,
    value %in% c("1.3_week" , "4.6_week") ~ 3,
    TRUE   ~ 4 )) %>% # "daily" | "several_daily"
  dplyr::select(-value) %>% 
  mutate(key = paste(key, "n", sep = "_")) %>% 
  spread(key = key, value = linear_value)
  
alcohol <- blood_data_summary_final %>% 
  dplyr::select(donor,beer, wine,liquor) %>% 
  gather(key = key, value =  value, -donor) %>% 
  mutate(linear_value = case_when(value %in% c( "never" , "very_rarely") ~ 1,
    value == "a_few_per_month" ~ 2,
    value == "a_few_per_week" ~ 3,
    TRUE ~ 4)) %>% # daily_or_almost
  dplyr::select(-value) %>% 
  mutate(key = paste(key, "n", sep = "_")) %>% 
  spread(key = key, value = linear_value)

blood_data_summary_final <- blood_data_summary_final %>% 
  full_join(food, by = "donor") %>% 
  full_join(alcohol, by = "donor") 

blood_data_summary_final <-
  blood_data_summary_final %>% 
  mutate(physical_act_daily_n = case_when(physical_act_daily == "under_15mn" ~ 0,
                                        physical_act_daily == "15_30mn" ~ 1,
                                        physical_act_daily == "30_60mn" ~ 2,
                                        TRUE ~ 3),
         physical_act_freq_n = case_when(physical_act_freq == "never" | 
                                           physical_act_freq == "less_than_monthly" ~ 0,
                                        physical_act_freq == "1_2_monthly" ~ 1,
                                        physical_act_freq == "1_weekly" ~ 2,
                                        TRUE ~ 3))
```

# Results - Population description

## Population description - Table 1


We recode the values for eating and drinking to make them interpretable. 

We code iron supplementation as a linear variable with a 4 grade scale:

   * 0: no supplementation
   * 1: less than half offered supplementation taken
   * 2: half or more offered supplementation taken
   * 4: All or nearly all offered supplementation taken

```{r}

# Recode food and non-alcoholic beverages
food <- blood_data_summary_final %>% 
  dplyr::select(donor,red_meat_n, fruit_berries_n,vegetables_n,
                fruit_juices_n,milk_n, coffee_n, tea_n ) %>% 
  gather(key = key, value =  value, -donor) %>% 
  mutate(value = case_when(value == 1 ~ "never",
    value == 2 ~ "less_than_once_weekly",
    value == 3 ~ "1_to_6_week",
    TRUE   ~ "daily_or_more" ),
    value = ordered(value, 
                    levels =  c( "never", "less_than_once_weekly", 
                               "1_to_6_week","daily_or_more"))) %>% 
  mutate(key = str_replace(key, "_n", "")) %>% 
  spread(key = key, value = value) 


alcohol <- blood_data_summary_final %>% 
  dplyr::select(donor,beer_n, wine_n,liquor_n) %>% 
  gather(key = key, value =  value, -donor) %>% 
 mutate(value = case_when(value == 1 ~ "rarely_or_never",
    value == 2 ~ "a_few_per_month",
    value == 3 ~ "a_few_per_week",
    TRUE   ~ "daily_or_almost" ),
    value = ordered(value, 
                    levels =  c( "rarely_or_never", "a_few_per_month", 
                               "a_few_per_week","daily_or_almost"))) %>% 
  mutate(key = str_replace(key, "_n", "")) %>% 
  spread(key = key, value = value) 




blood_data_summary_tbl <- blood_data_summary_final %>% 
  dplyr::select(donor,group, age,TwoYearsFromStartCount_FB,DaysToPreviousFB, Hb_v,
       Ferritin, TransferrinR,  CRP, DaysToPreviousFB, iron_comp_c, BMI,smoking,pregnancy,
       physical_act_freq, physical_act_daily) %>%
  full_join(food, by = "donor") %>%
  full_join(alcohol, by = "donor") %>%
    mutate(iron_supplementation = case_when(
    iron_comp_c == 3 ~ "All or almost all",
    iron_comp_c == 2 ~ "About half",
    iron_comp_c == 1 ~ "Less than half",
    TRUE ~ "None"),
    iron_supplementation = ordered(iron_supplementation, 
                  levels = c("None",   "Less than half", "About half","All or almost all")),
    physical_act_freq = case_when(physical_act_freq == "never" | 
                                    physical_act_freq == "less_than_monthly" ~ "less_than_monthly",
                                        physical_act_freq == "1_2_monthly" ~ "1_2_monthly",
                                        physical_act_freq == "1_weekly" ~ "1_weekly",
                                        TRUE ~ "twice_weekly_or_more"),
    physical_act_freq = ordered(physical_act_freq, 
                                levels = c("less_than_monthly","1_2_monthly", 
                                            "1_weekly","twice_weekly_or_more"))) 
    
```




```{r}
myVars <- c( "age", "BMI" ,"smoking", "pregnancy", "Hb_v",  "Ferritin", "TransferrinR",  "CRP",
             "TwoYearsFromStartCount_FB",  "DaysToPreviousFB",  "iron_supplementation", "red_meat", 
             "vegetables",  "fruit_berries", "milk",  "fruit_juices", "coffee", "tea", "beer", 
             "wine", "liquor", "physical_act_daily",  "physical_act_freq")

non_normal_vars <- c("DaysToPreviousFB", "Ferritin", "TransferrinR", "CRP", "BMI")

summary_table <- CreateTableOne(data = blood_data_summary_tbl %>% 
                                      mutate(group = dplyr::recode(group, 
                                                     Pre_menopause_women = "Pre-menopausal women", 
                                                     Post_menopause_women = "Post-menopausal women")) ,
                                vars = myVars, strata = "group",test = FALSE)

```

```{r}

tab3Mat <- print(summary_table, nonnormal = non_normal_vars, quote = FALSE, 
                 noSpaces = TRUE, printToggle = FALSE)

## Save to a CSV file
write.csv(tab3Mat, file = "../results/tables/Table_1_population.csv")

tab3Mat %>% 
  kable()

```

- 

## Paired group differences


We test for significant differences between group pairs with the relevant tests:

+ Independent t-tests for normally distributed variables
+ Mann-Whitney U test for non-normally distributed variables
+ $\chi^2$ test for categorical variables


```{r, warning= FALSE, message=FALSE}
# Test for normally distributed variables

get_t_test_norm <- function(data, excluded_group){
  
data %>% 
  dplyr::select(donor, group,age, Hb_v, TwoYearsFromStartCount_FB ) %>%
  filter(group != excluded_group) %>% 
  gather(key = variable, value = value, -donor, -group) %>% 
  droplevels() %>% 
  group_by(variable) %>% 
  do(tidy(t.test(value~group, var.equal = TRUE, data = .))) %>% 
  dplyr::select(variable, statistic, p.value, method) %>% 
  mutate(contrast = case_when(
    excluded_group == "Men" ~ "pre-menop vs post-menop",
    excluded_group == "Post_menopause_women" ~ "pre-menop vs men",
    TRUE ~ "post-menop vs men" )) -> result
    
    return(result)
}

for (excluded_group in levels(blood_data_summary_final$group)){
  if(!exists("t_test_norm")){
    t_test_norm <-  get_t_test_norm(blood_data_summary_tbl, excluded_group )
  }else{
    t_test_norm <-bind_rows(t_test_norm,  get_t_test_norm(blood_data_summary_tbl, excluded_group ))
  }
 
}

# Test for non-normally distributed variables

get_mann_Whit_test_norm <- function(data, excluded_group){
  
data %>% 
  dplyr::select(donor, group,DaysToPreviousFB, Ferritin, CRP, TransferrinR, BMI ) %>%
  filter(group != excluded_group) %>% 
  gather(key = variable, value = value, -donor, -group) %>% 
  droplevels() %>% 
  group_by(variable) %>% 
  do(tidy(wilcox.test(value~group,  paired = FALSE, data = .))) %>% 
  dplyr::select(variable, statistic, p.value, method) %>% 
  mutate(contrast = case_when(
    excluded_group == "Men" ~ "pre-menop vs post-menop",
    excluded_group == "Post_menopause_women" ~ "pre-menop vs men",
    TRUE ~ "post-menop vs men" )) -> result
    
    return(result)
}

for (excluded_group in levels(blood_data_summary_final$group)){
  if(!exists("t_test_non_norm")){
    t_test_non_norm <-  get_mann_Whit_test_norm(blood_data_summary_final,excluded_group )
  }else{
    t_test_non_norm <-bind_rows(t_test_non_norm,  
                                get_mann_Whit_test_norm(blood_data_summary_final,excluded_group ))
  }
 
}


# Test for categorical variables

get_chisqr_test_cat <- function(data, excluded_group){
  
  # Recode food and non-alcoholic beverages
  food <- data %>% 
    dplyr::select(donor,red_meat, fruit_berries,vegetables,fruit_juices,milk, coffee, tea ) %>% 
    gather(key = key, value =  value, -donor) %>% 
    mutate(linear_value = case_when(value == "never" ~ "1",
      value == "less_than_once_weekly" ~ "2",
      value %in% c("1.3_week" , "4.6_week") ~ "3",
      TRUE   ~ "4" )) %>% # "daily" | "several_daily"
    dplyr::select(-value) %>% 
    spread(key = key, value = linear_value)
    
  
  alcohol <- data %>% 
    dplyr::select(donor,beer, wine,liquor) %>% 
    gather(key = key, value =  value, -donor) %>% 
    mutate(linear_value = case_when(value %in% c( "never" , "very_rarely") ~ "1",
      value == "a_few_per_month" ~ "2",
      value == "a_few_per_week" ~ "3",
      TRUE ~ "4")) %>% # daily_or_almost
    dplyr::select(-value) %>% 
    spread(key = key, value = linear_value)
  
  
  data %>%
    mutate(iron_comp_f = case_when(
      iron_comp_c == 3 ~ "All or almost all",
      iron_comp_c == 2 ~ "About half",
      iron_comp_c == 1 ~ "Less than half",
      iron_comp_c == 0 ~ "None",
      TRUE ~ "NA")) %>% 
    dplyr::select(donor,group, smoking,iron_comp_f, physical_act_freq, physical_act_daily) %>%  
    mutate(physical_act_freq = case_when(physical_act_freq == "never" | 
                                        physical_act_freq == "less_than_monthly" ~ "less_than_monthly",
                                        physical_act_freq == "1_2_monthly" ~ "1_2_monthly",
                                        physical_act_freq == "1_weekly" ~ "1_weekly",
                                        TRUE ~ "twice_weekly_or_more"),
    physical_act_freq = ordered(physical_act_freq, 
                  levels = c("less_than_monthly", "1_2_monthly", 
                             "1_weekly","twice_weekly_or_more"))) %>% 
    full_join(food, by = "donor") %>% 
    full_join(alcohol, by = "donor") %>% 
    filter(group != excluded_group) %>% 
    gather(key = variable, value = value, -donor, -group) %>% 
    droplevels() %>% 
    group_by(variable, group, value) %>% 
    summarise(Freq = n()) %>% 
    ungroup() %>% 
    group_by(variable) %>% 
    do(tidy(chisq.test(xtabs(Freq ~ group + value,  data = .), simulate.p.value = FALSE))) %>% 
    dplyr::select(variable, statistic, p.value, method) -> result 
    
  if(excluded_group == "Men"){
    
    data %>%
    filter(group != excluded_group) %>% 
      droplevels() %>% 
    dplyr::select(donor,group, pregnancy) %>% 
    group_by( group, pregnancy) %>% 
    summarise(Freq = n()) %>% 
    ungroup() %>% 
    do(tidy(chisq.test(xtabs(Freq ~ group + pregnancy,  data = .)))) %>% 
    dplyr::select(-parameter) %>% 
    mutate(variable = "pregnancy") %>% 
    bind_rows(result) -> result
  } 
  
  
  
   result %>% 
      mutate(contrast = case_when(
        excluded_group == "Men" ~ "pre-menop vs post-menop",
        excluded_group == "Post_menopause_women" ~ "pre-menop vs men",
        TRUE ~ "post-menop vs men" )) -> result
   
   return(result)
}


for (excluded_group in levels(blood_data_summary_final$group)){
  if(!exists("chi_sqr_test")){
    chi_sqr_test <-  get_chisqr_test_cat(blood_data_summary_final,excluded_group )
  }else{
    chi_sqr_test <-bind_rows(chi_sqr_test,
                             get_chisqr_test_cat(blood_data_summary_final,excluded_group ))
  }
 
}

# chi_sqr_test

```


```{r, warning= FALSE, message=FALSE}
all_stat_test <- bind_rows(t_test_norm, t_test_non_norm, chi_sqr_test)

## Apply holm- bonferroni p correction
all_stat_test$p.value.adj <- p.adjust(all_stat_test$p.value, method = "holm") 

final_p_values <- all_stat_test %>% 
  dplyr::select( -statistic, -p.value) %>% 
    mutate(p.value.adj.final = case_when(p.value.adj < .0001 ~ "p < .0001",
                                       p.value.adj < .001 ~ "p < .001",
                                       p.value.adj < .01 ~ "p < .01",
                                       p.value.adj < .05 ~ "p < .05",
                                       TRUE ~ "n.s.")) %>% 
  dplyr::select( -p.value.adj) %>% 
  spread(key = contrast, value = p.value.adj.final) %>% 
  dplyr::select(-method) %>% 
  inner_join(all_stat_test %>% dplyr::select(variable, method) %>% distinct) %>% 
  ungroup() %>% 
  mutate(variable = dplyr::recode(variable, Hb_v = "Hemoglobin",
                           TwoYearsFromStartCount_FB = "Nb donations in last 2 years",
                           TransferrinR = "sTfR",
                           DaysToPreviousFB = "Nb of days since last donation",
                           iron_comp_f = "Iron supplementation",
                           red_meat = "Red meat",
                           fruit_berries = "Fruit and berries",
                           fruit_juices = "Fruit juices",
                           physical_act_freq = "Exercise frequency",
                           physical_act_daily = "Daily physical activity"),
    variable = ordered(variable, levels = c("age",  "BMI" ,"smoking", "pregnancy",
                                                 "Hemoglobin",  "Ferritin",  "CRP",  "sTfR",
                              "Nb donations in last 2 years", "Nb of days since last donation",    
             "Iron supplementation", "Red meat", "vegetables",  "Fruit and berries", "milk" , 
            "Fruit juices", "coffee", "tea", "beer",  "wine", "liquor", "Exercise frequency", 
            "Daily physical activity"))) %>% 
  arrange(variable) %>% 
  mutate(variable = as.character(variable))   

```



```{r warning=FALSE, results='asis'}

# Saving resuls of statistical tests to html file

stargazer(final_p_values, type = "html", 
          summary = FALSE, 
          rownames = FALSE,
          header = FALSE,
          out = "../results/tables/table_1_p_values_paired.html",
          initial.zero = FALSE, 
          digits = 5)
```

- 

# Results - Distributions of hemoglobin, ferritin and sTfR

## Incidence of anemia and iron deficiency

Unless otherwise specified, the thresholds used are the following:

* Hb: 117 g/L for women, 134 g/L for men ([Reference](https://www.terveyskirjasto.fi/terveyskirjasto/tk.koti?p_artikkeli=snk03031))
* Ferritin: 15 for all groups
* sTfR: 4.4 mg/L for women and 5 mg/L for men ([Reference](https://huslab.fi/cgi-bin/ohjekirja/tt_show.exe?assay=4720&terms=trans))
    
    
```{r}
# Create a DF with the thresholds for each group. 

threshold <- data.frame(group = c("Men", "Pre_menopause_women", "Post_menopause_women"), 
                       CRP_thresh = c(3, 3,3),
                       Hb_v_thresh = c(134, 117,117),
                       Ferritin_thresh = c(15, 15,15),
                       TransferrinR_thresh = c(5, 4.4, 4.4)) %>% 
  mutate(group = ordered(group, levels =  c("Pre_menopause_women", "Post_menopause_women", "Men"))) 

```

-

### Overall incidence of low hemoglobin, low ferritin and high sTfR

```{r message=FALSE, warning=FALSE}
# Identify donors with abnormal values
blood_data_summary_final <- blood_data_summary_final %>% 
  left_join(threshold, by="group") %>%
  mutate(low.fer = ifelse(Ferritin < Ferritin_thresh, TRUE, FALSE),
         anemia = ifelse(Hb_v < Hb_v_thresh, TRUE, FALSE),
         high.TfR = ifelse(TransferrinR > TransferrinR_thresh, TRUE, FALSE),
       logFerritin = log(Ferritin)) %>% 
  dplyr::select(-Ferritin_thresh, -Hb_v_thresh, -TransferrinR_thresh, -CRP_thresh)

threshold2 <- 
    threshold %>% 
    mutate(group = case_when(group == "Pre_menopause_women" ~ "Pre-menopausal women",
                             group == "Post_menopause_women" ~ "Post-menopausal women",
                             TRUE ~ "Men"),
         group = ordered(group),
         group = fct_rev(group)) 
  
new_donors_data <- new_donors_data %>% 
  left_join(threshold2, by="group") %>%
  mutate(low.fer = ifelse(Ferritin < Ferritin_thresh, TRUE, FALSE),
         anemia = ifelse(Hb_v < Hb_v_thresh, TRUE, FALSE),
         high.TfR = ifelse(TransferrinR > TransferrinR_thresh, TRUE, FALSE),
       logFerritin = log(Ferritin)) %>% 
  dplyr::select(-Ferritin_thresh, -Hb_v_thresh, -TransferrinR_thresh, -CRP_thresh)
```

The following table presents the proportion and 95%CI of donors with low hemoglobin, low ferritin or high sTfR for each demographic group.

```{r}
data <- blood_data_summary_final %>% 
  dplyr::select(donor, group, low.fer, anemia, high.TfR) %>% 
  mutate(group = dplyr::recode(group, Pre_menopause_women = "Pre-menopausal women",
         Post_menopause_women = "Post-menopausal women")) %>% 
  gather(key = key, value = value, -donor, -group) %>% 
  group_by(group, key) %>% 
  summarise(N_key = sum(value), N = n()) 

data %>% 
  bind_cols(data.frame(binconf(data$N_key, data$N))) %>% 
  mutate(prev = paste(round(PointEst,3)*100,
                      " (", round(Lower,3)*100,"-", 
                      round(Upper,3)*100, ")", sep = "")) %>% 
  dplyr::select(group, key, prev) %>% 
  spread(key = group, value = prev) %>% 
  kable()

```

-

### Overlap of low ferritin and high sTfR 


We compute the proportion (and the 95%CI) of donors who have low ferritin only, high sTfR only and both low ferritin and high sTfR. 


```{r message=FALSE, warning=FALSE}
blood_data_summary_final <- 
  blood_data_summary_final %>% 
    mutate(ID_only = low.fer & !high.TfR,
           ID_high_sTRF = low.fer & high.TfR,
           high_sTRF_only = high.TfR & !low.fer,
           either_or = low.fer | high.TfR) 

data <- blood_data_summary_final %>% 
  dplyr::select(donor, group, ID_only, ID_high_sTRF, high_sTRF_only) %>% 
  gather(key = key, value = value, -donor, -group) %>% 
  group_by(group, key) %>% 
  summarise(N_key = sum(value), N = n()) 

data %>% 
  bind_cols(data.frame(binconf(data$N_key, data$N))) %>% 
  mutate(prev = paste(round(PointEst,3)*100,
                      " (", round(Lower,3)*100,"-", 
                      round(Upper,3)*100, ")", sep = "")) %>% 
  ungroup() %>% 
  dplyr::select(group, key, prev) %>% 
  mutate(group = dplyr::recode(group, Pre_menopause_women = "Pre-menopausal women",
                                      Post_menopause_women = "Post-menopausal women"),
         key = dplyr::recode(key, ID_only = "Low ferritin only",
                                  high_sTRF_only = "High sTfR only", 
                                  ID_high_sTRF = "Both")) %>% 
  spread(key = group, value = prev) %>% 
  kable()

```

-

### Overlap between anemia (low hemoglobin) and iron deficiency


Here we look at the proportion of donors with low hemoglobin that also exhibit iron deficiency ( coded as ID: low ferritin and/or high sTfR).


```{r}
sum_data <-
  blood_data_summary_final %>% 
  filter(anemia) %>% 
     mutate(low_fer_only = low.fer & !high.TfR,
           low_fer_high_sTRF = low.fer & high.TfR,
           high_sTRF_only = high.TfR & !low.fer,
           ID = low.fer | high.TfR) %>% 
          mutate(group = dplyr::recode(group, Pre_menopause_women = "Pre-menopausal women",
                                      Post_menopause_women = "Post-menopausal women")) %>% 
         rename( Overall.iron.deficiency = ID,
                 Low.ferritin.only = ID_only,
                 High.sTfR.only = high_sTRF_only, 
                 Both = ID_high_sTRF) 


myVars <- c( "Overall.iron.deficiency", "Low.ferritin.only", "High.sTfR.only", "Both")

summary_table <- CreateTableOne(data = sum_data,  vars = myVars, strata = "group" )

tab3Mat <- print(summary_table, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)

tab3Mat %>% 
  kable()

```

-

### Further computations for discussion


First we compute, the proportion of donors identified as iron deficient (either low ferritin or high sTfR) that have high sTfR but normal ferritin for a ferritin threshold of 15 $\mu g/L$.


```{r}
myVars <- c("Low.ferritin.only", "High.sTfR.only", "Both")


summary_table <- 
 blood_data_summary_final %>% 
  mutate(low.fer =  Ferritin < 15) %>% 
    mutate(low_ferritin_only = low.fer & !high.TfR,
           both = low.fer & high.TfR,
           high_sTRF_only = high.TfR & !low.fer,
           either_or = low.fer | high.TfR) %>% 
    filter(either_or) %>%
    mutate(group = dplyr::recode(group, Pre_menopause_women = "Pre-menopausal women",
                                      Post_menopause_women = "Post-menopausal women")) %>% 
    rename(Low.ferritin.only = low_ferritin_only,
           High.sTfR.only = high_sTRF_only, 
           Both = both) %>% 
    CreateTableOne( vars = myVars, strata = "group" )

 tab3Mat <- print(summary_table, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)

tab3Mat %>% 
  kable()

```

Then we compute, the same proportion of donors identified as iron deficient (either low ferritin or high sTfR) that have high sTfR but normal ferritin for a ferritin threshold of 30 $\mu g/L$.


```{r}
summary_table <- 
 blood_data_summary_final %>% 
  mutate(low.fer =  Ferritin < 30) %>% 
    mutate(low_ferritin_only = low.fer & !high.TfR,
           both = low.fer & high.TfR,
           high_sTRF_only = high.TfR & !low.fer,
           either_or = low.fer | high.TfR) %>% 
    filter(either_or) %>% 
      mutate(group = dplyr::recode(group, Pre_menopause_women = "Pre-menopausal women",
                                      Post_menopause_women = "Post-menopausal women")) %>% 
    rename(Low.ferritin.only = low_ferritin_only,
           High.sTfR.only = high_sTRF_only, 
           Both = both) %>% 
  CreateTableOne( vars = myVars, strata = "group" )

 tab3Mat <- print(summary_table, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)

tab3Mat %>% 
  kable()
```



Finally, for the same 30 $\mu g/L$ ferritin threshold, we compute the proportion of donors with high sTfR who also have low ferritin.

```{r}
blood_data_summary_final %>% 
    mutate(low.fer =  Ferritin < 30) %>% 
  filter(high.TfR) %>% 
  mutate(group = dplyr::recode(group, Pre_menopause_women = "Pre-menopausal women",
                                Post_menopause_women = "Post-menopausal women")) %>% 
  group_by(group) %>% 
  summarise(Proportion.low.ferritin =mean(low.fer)) %>% 
  kable()
```

Proportion of low ferritin  donors who also have high sTfR: 

```{r}
blood_data_summary_final %>% 
    mutate(low.fer =  Ferritin <= 30) %>% 
  filter(low.fer) %>% 
  mutate(group = dplyr::recode(group, Pre_menopause_women = "Pre-menopausal women",
                              Post_menopause_women = "Post-menopausal women")) %>% 
  group_by(group) %>% 
  summarise(Proportion.High.sTfR =mean(high.TfR)) %>% 
  kable()
```

-

### CRP levels for anemic donors with and without ID


Here we check that high ferritin values for anemic donors was not a consequence of inflammation (high CRP). Results show that only one anemic donor has elevated CRP which could indicate inflammation. 

```{r}
blood_data_summary_final %>% 
  filter(anemia) %>% 
     mutate(ID_only = low.fer & !high.TfR,
           ID_high_sTRF = low.fer & high.TfR,
           high_sTRF_only = high.TfR & !low.fer,
           either_or = low.fer | high.TfR) %>% 
    mutate(group = dplyr::recode(group, Pre_menopause_women = "Pre-menopausal \n women",
         Post_menopause_women = "Post-menopausal \n women")) %>% 
  ggplot(aes(x = group, y = CRP, color = low.fer)) +
  geom_jitter(height = 0.05, width = 0.25, shape = 1) + 
  scale_color_discrete(name="Low\nferritin") +
   scale_y_log10()
```

## Figure 1

We plot sTfR as a function of ferritin, with color representing hemoglobin. Red lines represent threshold values for iron deficiency. Red circles indicate anemic donors whilst yellow crosses represent first-time donors who were not included in the analyses. 



```{r message=FALSE, warning=FALSE}
point_size = 1
  low_ferritin <- min(blood_data_summary_final$Ferritin)
  high_ferritin <- max(blood_data_summary_final$Ferritin)
  
  low_Hb <- min(blood_data_summary_final$Hb_v)
  high_Hb<- max(blood_data_summary_final$Hb_v)
  
  low_sTRF <-min(blood_data_summary_final$TransferrinR, na.rm = TRUE)
  high_sTRF<- max(blood_data_summary_final$TransferrinR, na.rm = TRUE)
  
  N_data <-
    blood_data_summary_final %>%
      group_by(group) %>% 
      summarise(N = n()) %>% 
      mutate(group_N = case_when(group == "Pre_menopause_women" ~ paste("Pre-menopausal women\n",
                                                                        "N = ",
                                                                        N,sep="" ),
                       group == "Post_menopause_women" ~ paste("Post-menopausal women\n","N = ",
                                                               N,sep="" ),
                       group == "Men" ~ paste("Men\n","N = ",N,sep="" ),
                       TRUE~ "NA")) 
  

  
   new_donors_data <- new_donors_data %>%
     mutate(group = fct_rev(group))

plt <- blood_data_summary_final %>%
  mutate(group = case_when(group == "Pre_menopause_women" ~ "Pre-menopausal women",
                   group == "Post_menopause_women" ~ "Post-menopausal women",
                   group == "Men" ~ "Men" ,
                   TRUE~ "NA"),
         group = ordered(group),
         group = fct_rev(group)) %>%
  ggplot(aes(x = Ferritin, y = TransferrinR)) +
  geom_vline(data = threshold2,
             aes(xintercept = Ferritin_thresh),
             color = "dark red",
             size = 0.3,
             linetype = 2) +
    geom_vline(xintercept = 30,
             color = "dark red",
             size = 0.3,
             linetype = 3) +
  geom_hline(data = threshold2,
             aes(yintercept = TransferrinR_thresh),
             color = "dark red",
             size = 0.3,
             linetype = 2) +
  geom_point(aes(color = Hb_v), size = point_size, alpha = 0.5) +
  geom_point(aes(alpha = anemia), shape = 1, color = "red", size = point_size*1.1 ) +
  geom_point(data = new_donors_data, shape = 8, size = point_size, color = "orange") +
  geom_point(data = new_donors_data, aes(alpha = anemia), shape = 1, color = "red", 
             size = point_size*1.5) +
  scale_color_viridis_c(direction = -1, option="D", trans = "log",
                        limits = c(low_Hb, high_Hb),
                        breaks=c(110, 117,134, 150, 175)) +
  scale_alpha_discrete(range = c(0,1), guide=FALSE) +
  scale_x_log10(limits = c(low_ferritin, high_ferritin)) +
  scale_y_log10(limits = c(low_sTRF, high_sTRF)) +
  theme_bw() +
  facet_grid(group ~ .) + 
               # labeller = label_wrap_gen(width = 2, multi_line = TRUE))  +
  theme(legend.position = c(0.15, 0.65), 
        panel.grid = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.border = element_rect(colour = NA),
        strip.background=element_rect(colour="white",fill="white"),
        strip.text = element_text(face="bold", size = 10),
        axis.line = element_line(colour="black"),
        axis.title = element_text(size = 10),
        # axis.title.y = element_blank(),
        axis.text = element_text(size = 9),
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 10),
        legend.background = element_blank()) +
    guides(color = guide_colorbar(barwidth = .75, barheight = 4, title = "Hb (g/l)")) +
  xlab(expression("Ferritin ("*mu* "g/l)")) + ylab("sTFR (mg/l)")# +

plt

 ggplot2::ggsave("../results/figures/fig_1.png", plt, height = 20, width = 8, units = "cm", dpi = 300)

```



```{r}

group_labels <- c(
  Pre_menopause_women = "Pre-menopausal women",
  Post_menopause_women = "Post-menopausal women",
  Men = "Men"
  )
blood_data_summary_final$group2 <- plyr::revalue(blood_data_summary_final$group,
  group_labels) 


threshold$group2 <- plyr::revalue(threshold$group,
  group_labels) 

```



## Population description - Iron supplementation

### Proportion of donors reporting being offered iron supplementation

First we investigate the proportion of donors who report being offered iron.

```{r out.width = '75%', fig.align= 'center'}
blood_data_summary_final %>% 
      mutate(group_N = case_when(group == "Pre_menopause_women" ~ "Pre-menopausal women" ,
                       group == "Post_menopause_women" ~ "Post-menopausal women" ,
                       group == "Men" ~ "Men" ,
                       TRUE~ "NA"),
            group_N = fct_rev(group_N)) %>% 
  group_by(group_N) %>% 
  summarise(proportion = mean(iron_supp)) %>% 
    ggplot(aes(x = group_N, y = proportion, fill= group_N)) +
  geom_col(position ="dodge") +
  geom_text(aes(label=round(proportion,2)), position=position_dodge(width=0.9), vjust=0.5) +
  scale_fill_manual(values = c( "#ff85a2",  "#de2d26","#00BFFF" ),
                     limits = c( "Pre-menopausal women", "Post-menopausal women", "Men" )) +
 theme(legend.position = "none", 
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = NA),
        strip.background=element_rect(colour="white",fill="white"),
        strip.text = element_text(face="bold", size = 10),
        axis.line = element_line(colour="black"),
        # axis.text.x=element_text(angle=20,hjust=1),
        axis.title.x =  element_text(size = 10),
       axis.title.y =  element_text(size = 10),
        axis.text = element_text(size = 10)) +
  # scale_y_log10(breaks=c(5, 15,25,50,100)) +
  xlab("Were you offered iron supplementation at your last visit?") +
  ylab("Proportion")

 
```


### Compliance when iron supplementation was provided

We then investigate how much iron was ingested by those donors who reported being offered iron. 

```{r}
blood_data_summary_final %>% 
  filter(iron_supp) %>% 
  mutate(iron_comp_f = case_when(iron_comp_c == 3 ~ "All or almost all",
    iron_comp_c == 2 ~ "About half",
    iron_comp_c == 1 ~ "Less than half",
    iron_comp_c == 0 ~ "None",
    TRUE ~ "NA"),
    iron_comp_f = ordered(iron_comp_f, 
                  levels = c("NA","None",   "Less than half", "About half","All or almost all"))) %>% 
     mutate(group_N = case_when(group == "Pre_menopause_women" ~ "Pre-menopausal women" ,
                       group == "Post_menopause_women" ~ "Post-menopausal women" ,
                       group == "Men" ~ "Men" ,
                       TRUE~ "NA"),
            group_N = fct_rev(group_N)) %>% 
  group_by(group_N, iron_comp_f) %>% 
  summarise(N = n()) %>% 
  ungroup() %>% 
  group_by(group_N) %>% 
  mutate(proportion = N/sum(N)) %>% 
  # gather(key = count_type, value = value, -group, -iron_comp_f) %>%
  ggplot(aes(x = iron_comp_f, y = proportion, fill= group_N)) +
  geom_col(position ="dodge") +
  geom_text(aes(label=round(proportion,2)), position=position_dodge(width=0.9), vjust=0.5) +
  scale_fill_manual(values = c( "#ff85a2",  "#de2d26","#00BFFF" ),
                     limits = c( "Pre-menopausal women", "Post-menopausal women", "Men" )) +
  facet_grid(.~ group_N, 
      labeller = label_wrap_gen(width = 2, multi_line = TRUE)) +
 theme(legend.position = "none", 
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = NA),
        strip.background=element_rect(colour="white",fill="white"),
        strip.text = element_text(face="bold", size = 10),
        axis.line = element_line(colour="black"),
        axis.text.x=element_text(angle=20,hjust=1),
        axis.title.x =  element_text(size = 8),
       axis.title.y =  element_text(size = 8),
        axis.text = element_text(size = 8)) +
  # scale_y_log10(breaks=c(5, 15,25,50,100)) +
  xlab("How much of the offered course of iron supplements did you take?") +
  ylab("Proportion")


```


# Results - Regression analyses

## Regressions on sTfR and ferritin  

We analyzed the data with multiple linear regression. We first ran Ordinary Least Square (OLS) regressions and their diagnostics. If the diagnostics identified problematic observations (e.g., outliers, influential observations), we also ran robust regression regressions [@yu2017robust]. Robust regression is based on less assumptions than OLS regression and is less sensitive to influential observations. It  gives more accurate estimates of regression coefficients in the presence of problematic observations. Using robust regression leads to a more principled approach to regression analysis and avoids potential confounds introduced by selective removal of influential observations. As the rlm command from the MASS package does not compute directly the *t* and *p* values, we used the f.robftest function from the sfsmisc package. Practical examples of using robust regression can be found in the literature [@davison2017platelet].  


In addition, we used relative importance analysis to estimate the average proportion of variance in the outcome variable explained by each co-variate [@Groemping2006]. Relative iportance was computed using the pmvd method which assigns 0 to the relative importance of a regressor if it is non significant in the complete model. .

The regressors were entered as follows in the regression models for sTfR and ferritin:

* Continuous variables were entered as continuous predictors
    * age, number of donations in last two years, nb of days since last donation, CRP, BMI
  
* Binary variables were entered as categorical (but coded as numerical for relative importance):
    * Smoking, pregnancy
  
* Ordinal variables were coded on a linear scale and entered as continuous [(Reference)](https://www3.nd.edu/~rwilliam/stats3/OrdinalIndependent.pdf)
    * Iron supplementation 
    * All dietary intake
  
We compute both coefficients and standardized coefficients (with all dependent and independent variables standardized) for robust regression and only coefficients for OLS regression. We present robust coefficients and their bootstrapped BCa 95% confidence intervals in Table 1 (ferritin) and  Table 2 (sTfR), robust standardized coefficients, their bootstrapped distribution and BCa 95% confidence intervals in Figure 2.A and non-standardized OLS regression coefficients in Supplementary Table 1 (Ferritin) and 2 (sTfR). 

## Pre-processing

### Rename and transform

```{r}
blood_data_summary_reg <- blood_data_summary_final %>% 
    rename(don_ct = TwoYearsFromStartCount_FB,
            last_don = DaysToPreviousFB,
            iron = iron_comp_c) %>% 
     mutate(don_ct_2 = don_ct^2,
            log_ferr = log(Ferritin),
            log_last_don = log(last_don)/log(2),
            log_CRP = log(CRP),
            log_sTfR = log(TransferrinR)) 
```

Data transformations:

* Log transformed variables: 
    * Ferritin
    * sTfR
    * CRP 
  
* Age is divided by 5 to simplify coeff interpretation
* Number of days before donation is transformed as $log\_last\_don = log(last\_don)/log(2)$ to help with the interpretation of coefficients. 
    * An increase in one of the transformed variable is equivalent to a doubling of the number of days since last donation.

* Standardization:
    * Standardized coefficients: 
        * All dependent and independent variables were standardized
      
    * Coefficients:
        * All dependent variables entered as continuous varaibles are centered but not scaled. 

## Pre-menopausal women
### Data pre-processing

We center all variables entered as continuous.

```{r}
# Center variables

test_data_std <- blood_data_summary_reg %>% 
  filter(group == "Pre_menopause_women") %>% 
  dplyr::select(donor ,red_meat_n, fruit_berries_n, vegetables_n, milk_n, 
                fruit_juices_n, tea_n, coffee_n, beer_n, wine_n, liquor_n) %>% 
  gather(key = key, value = value, -donor) %>% 
  group_by(key) %>% 
  mutate(value = scale(value, scale = FALSE)[,1]) %>% 
  spread(key = key, value = value)
  
test_data_std <- blood_data_summary_reg %>% 
  filter(group == "Pre_menopause_women") %>% 
  dplyr::select(donor, log_ferr, TransferrinR, age, log_CRP, BMI, don_ct,log_last_don,
                smoking, pregnancy, iron, log_sTfR) %>%
  mutate(age = age / 5, 
         age = scale(age, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         don_ct = scale(don_ct, scale = FALSE)[,1],
         log_last_don = scale(log_last_don, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         iron = scale(iron, scale = FALSE)[,1],
         don_ct_2 = don_ct^2) %>% 
  full_join(test_data_std, by = "donor")  %>% 
  dplyr::select(-donor)

```


### Mutiple regression - ferritin as outcome

#### Correlograms

```{r}

ggpairs(test_data_std, 
        columns = c("log_ferr", "age", "log_CRP",  "BMI", "don_ct", "don_ct_2",
                    "log_last_don", "smoking","iron"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```

```{r}

ggpairs(test_data_std, 
        columns = c("log_ferr", "pregnancy",  "red_meat_n", "vegetables_n", 
                    "fruit_berries_n", "milk_n", "fruit_juices_n"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```


```{r}

ggpairs(test_data_std, 
        columns = c("log_ferr", "coffee_n", "tea_n", "beer_n", "wine_n", "liquor_n"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```

#### OLS 

```{r }

lm7_pre_menop_std <-lm(log_ferr ~ age + BMI + log_CRP + smoking + pregnancy  + don_ct + don_ct_2 + 
                         log_last_don + iron + red_meat_n + vegetables_n + fruit_berries_n + milk_n +
                         fruit_juices_n +  coffee_n +  tea_n + beer_n + wine_n + liquor_n, test_data_std )

```

##### Model diagnostics

We create regression diagnostic plots in order to identify putative influential observations (see [Understanding Diagnostic Plots for Linear Regression Analysis](http://data.library.virginia.edu/diagnostic-plots/)).

 
```{r, message=FALSE, warning = FALSE}
test_data_std %>%
  rowid_to_column("ID") %>%
  mutate(outlier = ifelse(ID == 547 | ID == 561 | ID == 297 | ID == 59 | ID == 723, 
                          TRUE, FALSE)) -> test_data_std

autoplot(lm7_pre_menop_std, data = test_data_std, which = c(1:6), ncol = 3,  label.n = 5,
         alpha = 0.7, shape = 1, colour = "outlier") + 
  scale_color_manual(values = c( "black",  "red") )+
  theme_bw() + theme(legend.position = "none") 
  
pre_menop_diag_plot <- autoplot(lm7_pre_menop_std, data = test_data_std, which = c(1:2,6), ncol = 1, 
         alpha = 0.7, shape = 1, colour = "outlier", label.alpha = 0, size = 0.5) + 
  scale_color_manual(values = c( "dark grey",  "red") )+
  theme_classic() + theme(legend.position = "none") 

ggplot2::ggsave("../results/figures/supp_fig_3_pre_menop.svg", pre_menop_diag_plot, 
                    height = 11.5 * 1.25, width = 11.5 * 1.25/3, units = "cm", dpi = 300)

  
pre_menop_diag_plot
```


These plots identify several potentially problematic observations, therefore we also compute robust regression. 

#### Robust regression model and p-values


```{r, warning =  FALSE}

# https://stats.stackexchange.com/questions/205614/p-values-and-significance-in-rlm-mass-package-r?utm_medium=organic&utm_source=google_rich_qa&utm_campaign=google_rich_qa

# Run model 
robust_mdl_pre_menop_w <- rlm(log_ferr ~ age + BMI + log_CRP + smoking + pregnancy  + don_ct + don_ct_2 + 
                                log_last_don + iron + red_meat_n + vegetables_n + fruit_berries_n + milk_n +
                                fruit_juices_n +  coffee_n +  tea_n + beer_n + wine_n + liquor_n, test_data_std, maxit = 40)

#Tidy output
tidy_mdl <- tidy(robust_mdl_pre_menop_w)
# %>% 
#   filter(term != "(Intercept)")

# Compute p-values
regressor_list <- tidy_mdl$term
p_values <- rep(0,length(regressor_list ))
for (i in c(1:length(regressor_list))){
  p_values[[i]]<-f.robftest(robust_mdl_pre_menop_w, var = regressor_list[[i]])[[7]]
}

mdl_pre_menop_w <- tidy_mdl %>% 
  full_join(data.frame(term = regressor_list, p_value=p_values),
                      by="term") %>% 
  mutate(significance = case_when(
    p_value < 0.0001 ~ "p < 0.0001", 
    p_value < 0.001 ~ "p < 0.001", 
    p_value < 0.01 ~ "p < 0.01",
    p_value < 0.05 ~ "p < 0.05",
    TRUE ~"n.s."
  )) %>% 
  dplyr::select(-std.error, -statistic) 

# Save model in tidy form for future use
tidy_model_w_pre <- tidy(robust_mdl_pre_menop_w)

```

#### Model outputs and interpretation

```{r  warning=FALSE, results='asis'}

stargazer( robust_mdl_pre_menop_w, lm7_pre_menop_std,
          title = "Multiple linear regressions - Pre-menopausal women",
          intercept.bottom = FALSE,
          type = "html",
          single.row = TRUE,
          omit = "Constant",
          ci = FALSE,
          header = FALSE,
          dep.var.caption   = "Ferritin (log)", 
          dep.var.labels.include = FALSE,
          model.names = FALSE,
          model.numbers = FALSE,
          column.labels = c("Robust", "OLS"),
          covariate.labels =  c("Age", "BMI", "CRP", "Smoking(yes)",  "Pregnancy(Yes)",
                       "Nb donations (2 years)","(Nb donations (2 years))²", 
                       "Days since last donation",  "Iron supplementation", "Red meat",
                       "Vegetables", "Fruit and Berries", "Milk",
                         "Fruit Juices", "Coffee","Tea", "Beer", "Wine" , "Liquor"),
          star.cutoffs = c(0.05, 0.01, 0.001),
          report=('vcp'))
          
```

Results from robust and OLS rergessions are similar. Ferritin levels are positively correlated with age, BMI, Days since last donation and negatively correlated with nb of donations in teh last two years. 


  * An increase in 5 years of age is associated with a `r round(exp(tidy_model_w_pre %>% filter(term == "age") %>% magrittr::extract2("estimate"))-1,3)*100` % increase in Ferritin levels.

  * A doubling of the number of days since the last donation is associated with a `r round(exp(tidy_model_w_pre %>% filter(term == "log_last_don") %>% magrittr::extract2("estimate"))-1,3)*100` % increase in Ferritin levels.

  * An additional donation in the previous 2 years is associated with a `r round(exp(tidy_model_w_pre %>% filter(term == "don_ct") %>% magrittr::extract2("estimate"))-1,3)*-100` % decrease in Ferritin levels.
  
#### Relative importance

The relative importance analyses are run only on regressors that are consistently significant in at least one demographic group.

```{r, warning = FALSE, message=FALSE}
test_data_2_std<- test_data_std %>% 
  mutate(smoking = ifelse(smoking == "no", 0,1)) 

lm7_num_std <- lm(log_ferr ~ age + BMI + don_ct + don_ct_2 + log_last_don + 
                    smoking + iron + red_meat_n +
                    wine_n + beer_n, test_data_2_std)

bootresult <- boot.relimp(lm7_num_std, b = 1000, type = c("pmvd", "last",  "first"), fixed = FALSE)

plotdata <- booteval.relimp(bootresult, typesel = c("pmvd"), level = 0.95)


plot_data_lm7_pre_menop <- data_frame(regressor = names(plotdata@first),
                        lmg = plotdata@pmvd,
                        lmg_inf = plotdata@pmvd.lower[1,],
                        lmg_sup = plotdata@pmvd.upper[1,]) 

```


### Mutiple regression - sTfR as outcome

##### Correlograms


```{r }

ggpairs(test_data_std, 
        columns = c("log_sTfR", "age", "log_CRP",  "BMI", "don_ct", "don_ct_2", "log_last_don", "smoking","iron"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```

#### OLS

```{r}
lm7_pre_menop_sTfR <-lm(log_sTfR ~ age + BMI + log_CRP + smoking + pregnancy  + don_ct + don_ct_2 + 
                          log_last_don + iron + red_meat_n + vegetables_n + fruit_berries_n + milk_n + 
                          fruit_juices_n +  coffee_n +  tea_n + beer_n + wine_n + liquor_n, 
                        test_data_std )

```



###### Model diagnostics 

```{r message=FALSE, warning=FALSE}
test_data_std %>%
  # rowid_to_column("ID") %>%
  mutate(outlier = ifelse(ID == 60 | ID == 537 | ID == 567 | ID == 814 , TRUE, FALSE)) -> test_data_std

autoplot(lm7_pre_menop_sTfR, data = test_data_std, which = c(1:6), ncol = 3, label.n = 5,
         shape = 1, alpha = 0.7, colour = "outlier") + 
  scale_color_manual(values = c( "black",  "red") )+
  theme_bw() + theme(legend.position = "none") 
  
pre_menop_diag_plot <- autoplot(lm7_pre_menop_sTfR, data = test_data_std, which = c(1:2,6), ncol = 1, 
         shape = 1, alpha = 0.7, colour = "outlier", label.alpha = 0, size = 0.5) + 
  scale_color_manual(values = c( "dark grey",  "red") )+
  theme_classic() + theme(legend.position = "none") 

ggplot2::ggsave("../results/figures/supp_fig_3_pre_menop_sTfR.svg", pre_menop_diag_plot, 
                    height = 11.5 * 1.25, width = 11.5 * 1.25/3, units = "cm", dpi = 300)

  
pre_menop_diag_plot
```

#### Robust regression model amd p-values


```{r, warning =  FALSE}

# https://stats.stackexchange.com/questions/205614/p-values-and-significance-in-rlm-mass-package-r?utm_medium=organic&utm_source=google_rich_qa&utm_campaign=google_rich_qa

# Run model 
robust_mdl_sTfR_pre_menop_w <- rlm(log_sTfR ~ age + BMI + log_CRP + smoking + pregnancy  + don_ct + 
                                     don_ct_2 +  log_last_don + iron + red_meat_n + vegetables_n + 
                                     fruit_berries_n + milk_n +  fruit_juices_n +  coffee_n +  
                                     tea_n + beer_n + wine_n + liquor_n, test_data_std, 
                                      maxit = 40)

#Tidy output
tidy_mdl <- tidy(robust_mdl_sTfR_pre_menop_w)

# Compute p-values
regressor_list <- tidy_mdl$term
p_values <- rep(0,length(regressor_list ))
for (i in c(1:length(regressor_list))){
  p_values[[i]]<-f.robftest(robust_mdl_sTfR_pre_menop_w, var = regressor_list[[i]])[[7]]
}

mdl_sTfR_pre_menop_w <- tidy_mdl %>% 
  full_join(data.frame(term = regressor_list, p_value=p_values),
                      by="term") %>% 
  mutate(significance = case_when(
    p_value < 0.0001 ~ "p < 0.0001", 
    p_value < 0.001 ~ "p < 0.001", 
    p_value < 0.01 ~ "p < 0.01",
    p_value < 0.05 ~ "p < 0.05",
    TRUE ~"n.s."
  )) %>% 
  dplyr::select(-std.error, -statistic) 

```

#### Model outputs and interpretation

```{r warning=FALSE, results='asis'}

stargazer(robust_mdl_sTfR_pre_menop_w, lm7_pre_menop_sTfR, 
          title = "Pre-menopausal women",
          intercept.bottom = FALSE,
          type = "html",
          single.row = TRUE,
          omit = "Constant",
          ci = FALSE,
          header = FALSE,
          dep.var.caption   = "sTfR", 
          dep.var.labels.include = FALSE,
          model.names = FALSE,
          model.numbers = FALSE,
          column.labels = c("Robust", "OLS"),
          covariate.labels =  c("Age", "BMI", "CRP", "Smoking(yes)",  "Pregnancy(Yes)",
                       "Nb donations (2 years)","(Nb donations (2 years))²", 
                       "Days since last donation",  "iron supplementation", "Red meat",
                       "Vegetables", "Fruit and Berries", "Milk", "Fruit Juices", 
                       "Coffee","Tea", "Beer", "Wine" , "Liquor"),
          star.cutoffs = c(0.05, 0.01, 0.001),
          report=('vcp'))
          
```

The explained R² is smaller than for Ferritin. Iron supplementation has no significant influence on sTfR levels. The number of donations in the previous two years is negatively correlated with sTfR levels whilst the time since last donation does not have a significant effect. 

#### Relative importance 


Relative importance analyses are run only on regressors that are consistently significant in at least one demographic group.

```{r, warning = FALSE, message=FALSE}
test_data_2_std<- test_data_std %>% 
  mutate(smoking = ifelse(smoking == "no", 0,1)) 

lm7_num_std <- lm(log_sTfR ~ age + BMI + don_ct + log_CRP +  log_last_don + smoking +
                    red_meat_n + wine_n + 
                    beer_n + milk_n, test_data_2_std)

bootresult <- boot.relimp(lm7_num_std, b = 1000, type = c("pmvd", "last",  "first"), fixed = FALSE)

plotdata <- booteval.relimp(bootresult, typesel = c("pmvd"), level = 0.95)


plot_data_lm7_pre_menop_sTfR <- data_frame(regressor = names(plotdata@first),
                        lmg = plotdata@pmvd,
                        lmg_inf = plotdata@pmvd.lower[1,],
                        lmg_sup = plotdata@pmvd.upper[1,]) 

```

## Post-menopausal women

### Data pre-processing

We center all variables entered as continuous.

```{r}
# Center and standardize variables

test_data_std <- blood_data_summary_reg %>% 
  filter(group == "Post_menopause_women") %>% 
  dplyr::select(donor ,red_meat_n, fruit_berries_n, vegetables_n, milk_n, 
                fruit_juices_n, tea_n, coffee_n, beer_n, wine_n, liquor_n) %>% 
  gather(key = key, value = value, -donor) %>% 
  group_by(key) %>% 
  mutate(value = scale(value, scale = FALSE)[,1]) %>% 
  spread(key = key, value = value)
  
test_data_std <- blood_data_summary_reg %>% 
  filter(group == "Post_menopause_women") %>% 
  dplyr::select(donor, log_ferr, log_sTfR, age, log_CRP, BMI, don_ct,log_last_don,smoking, pregnancy, iron) %>% 
  mutate(age = age / 5, 
         age = scale(age, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         don_ct = scale(don_ct, scale = FALSE)[,1],
         log_last_don = scale(log_last_don, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         iron = scale(iron, scale = FALSE)[,1],
         don_ct_2 = don_ct^2) %>% 
  full_join(test_data_std, by = "donor")  %>% 
  dplyr::select(-donor)

```




### Mutiple regression - ferritin as outcome

#### Correlograms

```{r}

ggpairs(test_data_std, 
        columns = c("log_ferr", "age", "log_CRP",  "BMI", "don_ct", "don_ct_2", 
                    "log_last_don", "smoking","iron"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```

```{r }

ggpairs(test_data_std, 
        columns = c("log_ferr", "pregnancy", "red_meat_n", "vegetables_n", 
                    "fruit_berries_n", "milk_n", "fruit_juices_n"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```


```{r }

ggpairs(test_data_std, 
        columns = c("log_ferr", "coffee_n", "tea_n", "beer_n", "wine_n", "liquor_n"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```

#### OLS

```{r }

lm7_post_menop_std <- lm(log_ferr ~ age + BMI + log_CRP + smoking + pregnancy  + don_ct + don_ct_2 + 
                           log_last_don + iron + red_meat_n + vegetables_n + fruit_berries_n + milk_n + 
                           fruit_juices_n +  coffee_n +  tea_n + beer_n + wine_n + liquor_n,
                         test_data_std )

```

##### Model diagnostics


```{r, warning=FALSE}
test_data_std %>%
  rowid_to_column("ID") %>%
  mutate(outlier = ifelse(ID == 48 | ID == 157 | ID == 430 | ID == 445 | ID == 448, 
                          TRUE, FALSE)) -> test_data_std

autoplot(lm7_post_menop_std, data = test_data_std, which = c(1:6), ncol = 3, 
          label.n = 5,
         shape = 1, alpha = 0.7, colour = "outlier") + 
  scale_color_manual(values = c( "black",  "red") )+
  theme_bw() + theme(legend.position = "none") 
  
post_menop_diag_plot <- autoplot(lm7_post_menop_std, data = test_data_std,
                                 which = c(1:2,6), ncol = 1,  label.n = 5,
         shape = 1, alpha = 0.7, colour = "outlier", label.alpha = 0, size = 0.5) + 
  scale_color_manual(values = c( "dark grey",  "red") )+
  theme_classic() + theme(legend.position = "none") 

ggplot2::ggsave("../results/figures/supp_fig_3_post_menop.svg", post_menop_diag_plot, 
                    height = 11.5 * 1.25, width = 11.5 * 1.25/3, units = "cm", dpi = 300)

  
post_menop_diag_plot
```

There are 4 observations that are seem to be problematic, so we also run robust regressions. 

#### Robust regression model and p-values

```{r, warning =  FALSE}

# Run model 
robust_mdl_post_menop_w <- rlm(log_ferr ~ age + BMI + log_CRP + smoking + pregnancy  + don_ct + don_ct_2 + 
                                 log_last_don + iron + red_meat_n + vegetables_n + fruit_berries_n + milk_n +
                                 fruit_juices_n +  coffee_n +  tea_n + beer_n + wine_n + liquor_n, test_data_std,
                               maxit = 40)

#Tidy output
tidy_model_w_post <- tidy(robust_mdl_post_menop_w)

# Compute p-values
regressor_list <- tidy_model_w_post$term
p_values <- rep(0,length(regressor_list ))
for (i in c(1:length(regressor_list))){
  p_values[[i]]<-f.robftest(robust_mdl_post_menop_w, var = regressor_list[[i]])[[7]]
}

mdl_post_menop_w <- tidy_model_w_post %>% 
  full_join(data.frame(term = regressor_list, p_value = p_values),
                      by="term") %>% 
  mutate(significance = case_when(
    p_value < 0.0001 ~ "p < 0.0001", 
    p_value < 0.001 ~ "p < 0.001", 
    p_value < 0.01 ~ "p < 0.01",
    p_value < 0.05 ~ "p < 0.05",
    TRUE ~"n.s."
  )) %>% 
  dplyr::select(-std.error, -statistic) 

```

#### Model outputs and interpretation

```{r warning=FALSE, results='asis'}

stargazer(robust_mdl_post_menop_w, lm7_post_menop_std, 
          title = "Post-menopausal women",
          intercept.bottom = FALSE,
          type = "html",
          single.row = TRUE,
          omit = "Constant",
          ci = FALSE,
          header = FALSE,
          dep.var.caption   = "Ferritin", 
          dep.var.labels.include = FALSE,
          model.names = FALSE,
          model.numbers = FALSE,
          column.labels = c("Robust", "OLS"),
          covariate.labels =  c("Age", "BMI", "CRP", "Smoking(yes)",  "Pregnancy(Yes)",
                       "Nb donations (2 years)","(Nb donations (2 years))²", "Days since last donation", 
                        "iron supplementation", "Red meat","Vegetables", "Fruit and Berries", "Milk",
                         "Fruit Juices", "Coffee","Tea", "Beer", "Wine" , "Liquor"),
          star.cutoffs = c(0.05, 0.01, 0.001),
          report=('vcp'))
          
```

Results from robust and OLS regressions differ slightly. In both models, ferritin levels are positively correlated with  BMI, nb of days since last donation, and wine consumption (which would not survive multiple correction).


  * A doubling of the number of days since the last donation is associated with a `r round(exp(tidy_model_w_post %>% filter(term == "log_last_don") %>% magrittr::extract2("estimate"))-1,3)*100` % increase in Ferritin levels.

#### Relative importance

```{r, warning = FALSE, message = FALSE }
test_data_2_std<- test_data_std %>% 
  mutate(smoking = ifelse(smoking == "no", 0,1),
         pregnancy = ifelse(pregnancy == "no", 0,1)) 

lm7_num_std <- lm(log_ferr~ age + BMI + don_ct + don_ct_2+ log_last_don + smoking + iron + 
                    red_meat_n + wine_n + beer_n, test_data_2_std)


bootresult <- boot.relimp(lm7_num_std, b = 1000, type = c("pmvd", "last",  "first"), fixed = FALSE)

plotdata <- booteval.relimp(bootresult, typesel = c("pmvd"), level = 0.95)


plot_data_lm7_post_menop <- data_frame(regressor = names(plotdata@first),
                        lmg = plotdata@pmvd,
                        lmg_inf = plotdata@pmvd.lower[1,],
                        lmg_sup = plotdata@pmvd.upper[1,]) 

```

### Mutiple regression - sTfR as outcome
#### Correlograms

```{r}

ggpairs(test_data_std, 
        columns = c("log_sTfR", "age", "log_CRP",  "BMI", "don_ct", "don_ct_2", "log_last_don", "smoking","iron"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```

#### OLS

```{r }

# we run the sTfR mdoels here to take advanatge of the pre-processing

lm7_post_menop_sTfR <-lm(log_sTfR ~ age + BMI + log_CRP + smoking + pregnancy  + don_ct + don_ct_2 + 
                           log_last_don + iron + red_meat_n + vegetables_n + fruit_berries_n + milk_n + 
                           fruit_juices_n +  coffee_n +  tea_n + beer_n + wine_n + liquor_n, 
                         test_data_std )

```

##### Model diagnostics 

```{r, warning=FALSE}
test_data_std %>%
  # rowid_to_column("ID") %>%
  mutate(outlier = ifelse(ID == 249 | ID == 174 | ID == 200 | ID == 217 | ID == 208, 
                          TRUE, FALSE)) -> test_data_std

autoplot(lm7_post_menop_sTfR, data = test_data_std, which = c(1:6), ncol = 3, 
          label.n = 5,
         shape = 1, alpha = 0.7, colour = "outlier") + 
  scale_color_manual(values = c( "black",  "red") )+
  theme_bw() + theme(legend.position = "none") 
  
post_menop_diag_plot <- autoplot(lm7_post_menop_sTfR, data = test_data_std, which = c(1:2,6), ncol = 1, 
         shape = 1, alpha = 0.7, colour = "outlier", label.alpha = 0, size = 0.5) + 
  scale_color_manual(values = c( "dark grey",  "red") )+
  theme_classic() + theme(legend.position = "none") 

ggplot2::ggsave("../results/figures/supp_fig_3_post_menop_sTfR.svg", post_menop_diag_plot, 
                    height = 11.5 * 1.25, width = 11.5 * 1.25/3, units = "cm", dpi = 300)

pre_menop_diag_plot
```


There are several problematic observations, so we also run robust regression. 

#### Robust regression model and p-values


```{r, warning =  FALSE}

# Run model 
robust_mdl_sTfR_post_menop_w <- rlm(log_sTfR ~ age + BMI + log_CRP + smoking + pregnancy  + don_ct + don_ct_2 + 
                                      log_last_don + iron + red_meat_n + vegetables_n + fruit_berries_n + milk_n + 
                                      fruit_juices_n +  coffee_n +  tea_n + beer_n + wine_n + liquor_n, test_data_std, 
                                    maxit = 40)

#Tidy output
tidy_mdl <- tidy(robust_mdl_sTfR_post_menop_w)

# Compute p-values
regressor_list <- tidy_mdl$term
p_values <- rep(0,length(regressor_list ))
for (i in c(1:length(regressor_list))){
  p_values[[i]]<-f.robftest(robust_mdl_sTfR_post_menop_w, var = regressor_list[[i]])[[7]]
}

mdl_sTfR_post_menop_w <- tidy_mdl %>% 
  full_join(data.frame(term = regressor_list, p_value=p_values),
                      by="term") %>% 
  mutate(significance = case_when(
    p_value < 0.0001 ~ "p < 0.0001", 
    p_value < 0.001 ~ "p < 0.001", 
    p_value < 0.01 ~ "p < 0.01",
    p_value < 0.05 ~ "p < 0.05",
    TRUE ~"n.s."
  )) %>% 
  dplyr::select(-std.error, -statistic) 

```

#### Model outputs and interpretation

```{r, warning=FALSE, results='asis'}
stargazer(robust_mdl_sTfR_post_menop_w, lm7_post_menop_sTfR, 
          title = "Multiple linear regressions - Post-menopausal women",
          intercept.bottom = FALSE,
          type = "html",
          single.row = TRUE,
          omit = "Constant",
          ci = FALSE,
          header = FALSE,
          dep.var.caption   = "sTfR", 
          dep.var.labels.include = FALSE,
          model.names = FALSE,
          model.numbers = FALSE,
          column.labels = c("Robust", "OLS"),
          covariate.labels =  c("Age", "BMI", "CRP", "Smoking(yes)",  "Pregnancy(Yes)",
                       "Nb donations (2 years)","(Nb donations (2 years))²", "Days since last donation", 
                        "iron supplementation", "Red meat","Vegetables", "Fruit and Berries", "Milk",
                         "Fruit Juices", "Coffee","Tea", "Beer", "Wine" , "Liquor"),
          star.cutoffs = c(0.05, 0.01, 0.001),
          report=('vcp'))
          
```


The regression R² is only .06, and the overall regression is not significant.  Our covariates cannot significantly explain inter-individual variability in sTfR. 

#### Relative importance of regressors 

This analysis is run only on regressors that are consistently significant in at least one demographic group. 


```{r , warning = FALSE}
test_data_2_std<- test_data_std %>% 
  mutate(smoking = ifelse(smoking == "no", 0,1)) 

lm7_num_std <- lm(log_sTfR ~ age + BMI + don_ct + log_CRP +  log_last_don + smoking  + 
                    red_meat_n + wine_n + beer_n + milk_n, 
                  test_data_2_std)

bootresult <- boot.relimp(lm7_num_std, b = 1000, type = c("pmvd", "last",  "first"), fixed = FALSE)

plotdata <- booteval.relimp(bootresult, typesel = c("pmvd"), level = 0.95)

plot_data_lm7_post_menop_sTfR <- data_frame(regressor = names(plotdata@first),
                        lmg = plotdata@pmvd,
                        lmg_inf = plotdata@pmvd.lower[1,],
                        lmg_sup = plotdata@pmvd.upper[1,]) 
```


## Men

### Data pre-processing

```{r}
# Center and standardize variables

test_data_std <- blood_data_summary_reg %>% 
  filter(group == "Men") %>% 
  dplyr::select(donor ,red_meat_n, fruit_berries_n, vegetables_n, milk_n, 
                fruit_juices_n, tea_n, coffee_n, beer_n, wine_n, liquor_n) %>% 
  gather(key = key, value = value, -donor) %>% 
  group_by(key) %>% 
  mutate(value = scale(value, scale = FALSE)[,1]) %>% 
  spread(key = key, value = value)
  
test_data_std <- blood_data_summary_reg %>% 
  filter(group == "Men") %>% 
  dplyr::select(donor, log_ferr,log_sTfR, age, log_CRP, BMI, don_ct,log_last_don,smoking, iron) %>% 
  mutate(age = age / 5, 
         age = scale(age, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         don_ct = scale(don_ct, scale = FALSE)[,1],
         log_last_don = scale(log_last_don, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         iron = scale(iron, scale = FALSE)[,1],
         don_ct_2 = don_ct^2) %>% 
  full_join(test_data_std, by = "donor")  %>% 
  dplyr::select(-donor)


```


### Mutiple regression - ferritin as outcome

#### Correlograms

```{r }

ggpairs(test_data_std, 
        columns = c("log_ferr", "age", "log_CRP",  "BMI", "don_ct", "don_ct_2", "log_last_don", 
                    "smoking","iron"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```


```{r }

ggpairs(test_data_std, 
        columns = c("log_ferr", "red_meat_n", "vegetables_n", "fruit_berries_n", 
                    "milk_n", "fruit_juices_n"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```


```{r }

ggpairs(test_data_std, 
        columns = c("log_ferr", "coffee_n", "tea_n", "beer_n", "wine_n", "liquor_n"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```

#### OLS

```{r }

lm7_men_std <-lm(log_ferr ~ age + BMI + log_CRP + smoking  + don_ct + don_ct_2 + 
                  log_last_don + iron + red_meat_n + vegetables_n + fruit_berries_n + 
                  milk_n + fruit_juices_n +  coffee_n +  tea_n + beer_n + wine_n + liquor_n, 
                 test_data_std )
```


##### Model diagnostics


```{r warning=FALSE, message=FALSE}
test_data_std %>%
  rowid_to_column("ID") %>%
  mutate(outlier = ifelse(ID == 433 | ID == 449 | ID == 583 | ID == 4 | ID == 158 | ID == 319, 
                          TRUE, FALSE)) -> test_data_std

autoplot(lm7_men_std, data = test_data_std, which = c(1:6), ncol = 3,  label.n = 5,
         shape = 1, alpha = 0.7, colour = "outlier") + 
  scale_color_manual(values = c( "black",  "red") )+
  theme_bw() + theme(legend.position = "none") 
  

  
  men_diag_plot <- autoplot(lm7_men_std, data = test_data_std, which = c(1:2,6), ncol = 1, 
         shape = 1, alpha = 0.7, colour = "outlier", label.alpha = 0, size = 0.3) + 
  scale_color_manual(values = c( "dark grey",  "red") )+
  theme_classic() + theme(legend.position = "none") 
  
  ggplot2::ggsave("../results/figures/supp_fig_3_men.svg", men_diag_plot, 
                    height = 11.5 * 1.25, width = 11.5 * 1.25/3, units = "cm", dpi = 300)

men_diag_plot
```


There are seevral problemetic observations, so we also run robust regressions. 

#### Robust regression model and p-values

```{r, warning =  FALSE}

# https://stats.stackexchange.com/questions/205614/p-values-and-significance-in-rlm-mass-package-r?utm_medium=organic&utm_source=google_rich_qa&utm_campaign=google_rich_qa

# Run model 
robust_mdl_men <- rlm(log_ferr ~ age + BMI + log_CRP + smoking + don_ct + don_ct_2 + 
                      log_last_don + iron + red_meat_n + vegetables_n + fruit_berries_n + 
                      milk_n + fruit_juices_n +  coffee_n +  tea_n + beer_n + wine_n + liquor_n, 
                      test_data_std, maxit = 40)

#Tidy output
tidy_model_men <- tidy(robust_mdl_men)

# Compute p-values
regressor_list <- tidy_model_men$term
p_values <- rep(0,length(regressor_list ))
for (i in c(1:length(regressor_list))){
  p_values[[i]]<-f.robftest(robust_mdl_men, var = regressor_list[[i]])[[7]]
}

mdl_men <- tidy_model_men %>% 
  full_join(data.frame(term = regressor_list, p_value=p_values),
                      by="term") %>% 
  mutate(significance = case_when(
    p_value < 0.0001 ~ "p < 0.0001", 
    p_value < 0.001 ~ "p < 0.001", 
    p_value < 0.01 ~ "p < 0.01",
    p_value < 0.05 ~ "p < 0.05",
    TRUE ~"n.s."
  )) 

```


#### Model outputs and interpretation

```{r , warning=FALSE, results='asis'}
stargazer(robust_mdl_men, lm7_men_std,
          title = "Multiple linear regressions - Men",
          intercept.bottom = FALSE,
          type = "html",
          single.row = TRUE,
          omit = "Constant",
          ci = FALSE,
          header = FALSE,
          dep.var.caption   = "Ferritin", 
          dep.var.labels.include = FALSE,
          model.names = FALSE,
          model.numbers = FALSE,
          column.labels = c("Robust", "OLS"),
          covariate.labels =  c("Age", "BMI", "CRP", "Smoking(yes)",
                       "Nb donations (2 years)","(Nb donations (2 years))²", "Days since last donation", 
                        "iron supplementation", "Red meat","Vegetables", "Fruit and Berries", "Milk",
                         "Fruit Juices", "Coffee","Tea", "Beer", "Wine" , "Liquor"),
          star.cutoffs = c(0.05, 0.01, 0.001),
          report=('vcp'))
          
```

Results from robust and OLS regressions are similar. In both models, ferritin levels are positively correlated with  age, BMI, nb of days since last donation,  red meat consumption as well as Beer and wine consumption (would not survive multiple comparison correction). and wine consumption (which would not survive multiple correction). Ferritin levels are negatively correlated with the number of donations and iron supplementation. A significant negative corelation with the nb of donations squared indicates that the effect of nb of donations is larger for small numbers of donations than for large number of donations.

  * An increase in 5 years of age is associated with a `r round(exp(tidy_model_men %>% filter(term == "age") %>% magrittr::extract2("estimate"))-1,3)*100` % increase in Ferritin levels.

  * A doubling of the number of days since the last donation is associated with a `r round(exp(tidy_model_men %>% filter(term == "log_last_don") %>% magrittr::extract2("estimate"))-1,3)*100` % increase in Ferritin levels.
  
  * An increase of meat consumption of one on our grading scale is associated with a `r round(exp(tidy_model_men %>% filter(term == "red_meat_n") %>% magrittr::extract2("estimate"))-1,3)*100` % increase in Ferritin levels.

  * An increase of iron supplementation of one on our grading scale is associated with a `r round(exp(tidy_model_men %>% filter(term == "iron") %>% magrittr::extract2("estimate"))-1,3)*-100` % decrease in Ferritin levels.
  
  * An additional donation in the previous 2 years is associated with a `r round(exp(tidy_model_men %>% filter(term == "don_ct") %>% magrittr::extract2("estimate"))-1,3)*-100` % decrease in Ferritin levels.

#### Relative importance


```{r , warning = FALSE}
test_data_2_std<- test_data_std %>% 
  mutate(smoking = ifelse(smoking == "no", 0,1)) 

lm7_num_std <- lm(log_ferr~ age + BMI + don_ct + don_ct_2+ log_last_don + smoking + 
                    iron + red_meat_n + wine_n + beer_n, test_data_2_std)

bootresult <- boot.relimp(lm7_num_std, b = 1000, type = c("pmvd", "last",  "first"), fixed = FALSE)

plotdata <- booteval.relimp(bootresult, typesel = c("pmvd"), level = 0.95)

plot_data_lm7_men <- data_frame(regressor = names(plotdata@first),
                        lmg = plotdata@pmvd,
                        lmg_inf = plotdata@pmvd.lower[1,],
                        lmg_sup = plotdata@pmvd.upper[1,]) 

```

### Mutiple regression - sTfR as outcome

#### Correlograms


```{r }

ggpairs(test_data_std, 
        columns = c("log_sTfR", "age", "log_CRP",  "BMI", "don_ct", "don_ct_2", 
                    "log_last_don", "smoking","iron"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```

#### OLS

```{r }

# we run the sTfR mdoels here to take advanatge of the pre-processing

lm7_men_sTfR <-lm(log_sTfR ~ age + BMI + log_CRP + smoking  + don_ct + don_ct_2 + 
                    log_last_don + iron + red_meat_n + vegetables_n + fruit_berries_n +
                    milk_n + fruit_juices_n +  coffee_n +  tea_n + beer_n + wine_n + liquor_n, 
                  test_data_std )

```

##### Model diagnostics 

```{r, warning=FALSE}
test_data_std %>%
   # rowid_to_column("ID") %>%
  mutate(outlier = ifelse(ID == 465 | ID == 91 | ID == 692 | ID == 342 | ID == 55 | ID == 339, 
                          TRUE, FALSE)) -> test_data_std

autoplot(lm7_men_sTfR, data = test_data_std, 
         which = c(1:6), 
         ncol = 3, 
         label.n = 5,
         shape = 1, alpha = 0.7, colour = "outlier") + 
  scale_color_manual(values = c( "black",  "red") )+
  theme_bw() + theme(legend.position = "none") 
  
post_menop_diag_plot <- autoplot(lm7_men_sTfR, data = test_data_std, which = c(1:2,6), ncol = 1, 
         shape = 1, alpha = 0.7, colour = "outlier", label.alpha = 0, size = 0.5) + 
  scale_color_manual(values = c( "dark grey",  "red") )+
  theme_classic() + theme(legend.position = "none") 

ggplot2::ggsave("../results/figures/supp_fig_3_men_sTfR.svg", post_menop_diag_plot, 
                    height = 11.5 * 1.25, width = 11.5 * 1.25/3, units = "cm", dpi = 300)

  
pre_menop_diag_plot
```

#### Robust regression model and p-values


```{r, warning =  FALSE}
# Run model 
robust_mdl_sTfR_men <- rlm(log_sTfR ~ age + BMI + log_CRP + smoking   + don_ct + don_ct_2 + log_last_don +
                             iron + red_meat_n + vegetables_n + fruit_berries_n + milk_n + fruit_juices_n + 
                             coffee_n +  tea_n + beer_n + wine_n + liquor_n, test_data_std, maxit = 40)

#Tidy output
tidy_mdl <- tidy(robust_mdl_sTfR_men)

# Compute p-values
regressor_list <- tidy_mdl$term
p_values <- rep(0,length(regressor_list ))
for (i in c(1:length(regressor_list))){
  p_values[[i]]<-f.robftest(robust_mdl_sTfR_men, var = regressor_list[[i]])[[7]]
}

mdl_sTfR_men <- tidy_mdl %>% 
  full_join(data.frame(term = regressor_list, p_value=p_values),
                      by="term") %>% 
  mutate(significance = case_when(
    p_value < 0.0001 ~ "p < 0.0001", 
    p_value < 0.001 ~ "p < 0.001", 
    p_value < 0.01 ~ "p < 0.01",
    p_value < 0.05 ~ "p < 0.05",
    TRUE ~"n.s."
  )) %>% 
  dplyr::select(-std.error, -statistic) 

```

#### Model outputs and interpretation

```{r , warning=FALSE, results='asis'}
stargazer(robust_mdl_sTfR_men, lm7_men_sTfR, 
          title = "Multiple regression - Men",
          intercept.bottom = FALSE,
          type = "html",
          single.row = TRUE,
          omit = "Constant",
          ci = FALSE,
          header = FALSE,
          dep.var.caption   = "sTfR", 
          dep.var.labels.include = FALSE,
          model.names = FALSE,
          model.numbers = FALSE,
          column.labels = c("Robust", "OLS"),
          covariate.labels =  c("Age", "BMI", "CRP", "Smoking(yes)",  
                       "Nb donations (2 years)","(Nb donations (2 years))²", "Days since last donation", 
                        "iron supplementation", "Red meat","Vegetables", "Fruit and Berries", "Milk",
                         "Fruit Juices", "Coffee","Tea", "Beer", "Wine" , "Liquor"),
          star.cutoffs = c(0.05, 0.01, 0.001),
          report=('vcp'))
          
```

The R² is smaller than the one obtained for ferritin for the same demographic group. Nb of donations in the last two years but not time since last donations significantly co-varies with sTfR.


#### Relative importance


```{r , warning = FALSE}
test_data_2_std<- test_data_std %>% 
  mutate(smoking = ifelse(smoking == "no", 0,1)) 

lm7_num_std <- lm(log_sTfR ~ age + BMI + don_ct + log_CRP +  log_last_don + smoking  + red_meat_n + 
                    wine_n + beer_n + milk_n, test_data_2_std)

bootresult <- boot.relimp(lm7_num_std, b = 1000, type = c("pmvd", "last",  "first"), fixed = FALSE)

plotdata <- booteval.relimp(bootresult, typesel = c("pmvd"), level = 0.95)

plot_data_lm7_men_sTfR <- data_frame(regressor = names(plotdata@first),
                        lmg = plotdata@pmvd,
                        lmg_inf = plotdata@pmvd.lower[1,],
                        lmg_sup = plotdata@pmvd.upper[1,]) 

```


## Regression Coefficient bootstrapped CIs

We compute bootstrapped distributions for regression coefficients for OLS and robust regressions as well as standardized regression coefficients for robust regressions. From these bootstrapped distributions we furthermore compute the 95% BCa confidence intervals.

### OLS regression - coefficients

```{r}

# Create tidy versions of OLS regression models

lm7_pre_menop_df_std <- tidy(lm7_pre_menop_std) %>% 
  mutate(beta_sign = sign(estimate),
         coefficient = estimate,
   group  = "Pre_menopause_women",
   is_sig = ifelse(p.value < .05, TRUE, FALSE)) %>% 
  dplyr::select(term, beta_sign, group,is_sig,coefficient)

lm7_post_menop_df_std <- tidy(lm7_post_menop_std) %>% 
  mutate(beta_sign = sign(estimate),
   group  = "Post_menopause_women", 
            coefficient = estimate,
   is_sig = ifelse(p.value < .05, TRUE, FALSE)) %>%  
  dplyr::select(term, beta_sign, group,is_sig,coefficient)

lm7_men_df_std <- tidy(lm7_men_std) %>% 
  mutate(beta_sign = sign(estimate),
   group  = "Men",
    coefficient = estimate,
   is_sig = ifelse(p.value < .05, TRUE, FALSE)) %>%  
  dplyr::select(term, beta_sign, group,is_sig,coefficient)

lm_signs_std <- bind_rows(lm7_pre_menop_df_std,lm7_post_menop_df_std,lm7_men_df_std) %>% 
  rename(regressor = term) %>% 
  filter(regressor != "(Intercept)") 

lm7_pre_menop_df_sTfR <- tidy(lm7_pre_menop_sTfR) %>% 
  mutate(beta_sign = sign(estimate),
         coefficient = estimate,
   group  = "Pre_menopause_women",
   is_sig = ifelse(p.value < .05, TRUE, FALSE)) %>% 
  dplyr::select(term, beta_sign, group,is_sig,coefficient)

lm7_post_menop_df_sTfR <- tidy(lm7_post_menop_sTfR) %>% 
  mutate(beta_sign = sign(estimate),
   group  = "Post_menopause_women", 
            coefficient = estimate,
   is_sig = ifelse(p.value < .05, TRUE, FALSE)) %>%  
  dplyr::select(term, beta_sign, group,is_sig,coefficient)

lm7_men_df_sTfR <- tidy(lm7_men_sTfR) %>% 
  mutate(beta_sign = sign(estimate),
   group  = "Men",
    coefficient = estimate,
   is_sig = ifelse(p.value < .05, TRUE, FALSE)) %>%  
  dplyr::select(term, beta_sign, group,is_sig,coefficient)

lm_signs_sTfR <- bind_rows(lm7_pre_menop_df_sTfR, lm7_post_menop_df_sTfR, lm7_men_df_sTfR) %>% 
  rename(regressor = term) %>% 
  filter(regressor != "(Intercept)") 
```


```{r}

# https://www.painblogr.org/2017-10-18-purrring-through-bootstraps.html


get_coefficients <- function(data, boot_ind){
  fit <- lm (log_ferr~.,data[boot_ind,])
  return(coef(fit))
}

get_coefficients_sTfR <- function(data, boot_ind){
  fit <- lm (log_sTfR~.,data[boot_ind,])
  return(coef(fit))
}

compute_Bca_CI <-function(fit_boot){
  Bca_inf = rep(0, length(fit_boot$t0))
  Bca_sup = rep(0, length(fit_boot$t0))
  for (i_regressor in 1:length(fit_boot$t0)){
    CI <- boot.ci(fit_boot, type = "bca", index=i_regressor)
    Bca_inf[i_regressor] <- CI$bca[4]
    Bca_sup[i_regressor] <- CI$bca[5]
  }
  return(tibble(Bca_inf,Bca_sup, regressor = names(fit_boot$t0)) %>% 
           filter(regressor != "(Intercept)"))
}


get_bootstrap <- function(data, nb_boot, outcome){
 
  if (outcome == "Ferritin"){
    data <- data %>% dplyr::select(-log_sTfR)
    
    fit_boot <- boot(data, strata = data$smoking, statistic= get_coefficients, R = nb_boot)
  }else{
    data <- data %>% dplyr::select(-log_ferr)
    
    fit_boot <- boot(data, strata = data$smoking, statistic= get_coefficients_sTfR, R = nb_boot)
  }

  fit_boot_distrib <- as.tibble(fit_boot$t)  
  names(fit_boot_distrib) <- names(fit_boot$t0)

  fit_boot_Bca <- compute_Bca_CI(fit_boot)
  
  return(list(fit_boot_distrib,fit_boot_Bca))
}




get_bootstrap_coeffs <- function(blood_data_summary_reg, current_group, outcome, nb_boot )
{

  ## preprocess and standardize data
  test_data_std <- blood_data_summary_reg %>% 
    filter(group == current_group) %>% 
    dplyr::select(donor ,red_meat_n, fruit_berries_n, vegetables_n, milk_n, fruit_juices_n, 
                  tea_n, coffee_n, beer_n, wine_n, liquor_n ) %>% 
    gather(key = key, value = value, -donor) %>% 
    group_by(key) %>% 
    mutate(value = scale(value, scale = FALSE)[,1]) %>% 
    spread(key = key, value = value)
  
  test_data_std <- blood_data_summary_reg %>% 
    filter(group == current_group) %>% 
    dplyr::select(donor, log_ferr, log_sTfR, age, log_CRP, BMI, don_ct,log_last_don,smoking, 
                  pregnancy, iron) %>% 
  mutate(age = age / 5, 
         age = scale(age, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         don_ct = scale(don_ct, scale = FALSE)[,1],
         log_last_don = scale(log_last_don, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         iron = scale(iron, scale = FALSE)[,1],
         don_ct_2 = don_ct^2) %>% 
    full_join(test_data_std, by = "donor")   %>% 
    dplyr::select(-donor) 
  
  if(current_group == "Men"){
    test_data_std <-  dplyr::select(test_data_std, -pregnancy)
  }

  
  result <- test_data_std %>% 
  get_bootstrap(nb_boot = nb_boot, outcome = outcome)


  bootstrap_distrib <- result[[1]] %>% 
    gather(key = regressor, value = coefficient) %>% 
    filter(regressor != "(Intercept)") %>% 
    mutate(group = current_group)
    
  Bca_CI  <- result[[2]] %>% mutate(group = current_group)

  return(list(bootstrap_distrib,Bca_CI))
}

set.seed(125)
output_file_distrib = ("../results/bootstraps/coeff_boot_distrib_seed_125_ctr_strata_final.rdata")
output_file_Bca = ("../results/bootstraps/coeff_boot_Bca_seed_125_ctr_strata_final.rdata")
# 
output_file_distrib_sTfR = ("../results/bootstraps/coeff_boot_distrib_seed_125_ctr_sTfR_final.rdata")
output_file_Bca_sTfR = ("../results/bootstraps/coeff_boot_Bca_seed_125_ctr_sTfR_final.rdata")


get_from_file_ferr = TRUE

if (get_from_file_ferr){
  load(file=output_file_distrib)
  load(file=output_file_Bca)
}else
{
   for(group_str in c("Pre_menopause_women", "Post_menopause_women","Men")){
    
    stuff <- get_bootstrap_coeffs(blood_data_summary_reg, group_str, outcome = "Ferritin", nb_boot = 10000)
    
    if(!exists("bootstrap_distrib")){
      bootstrap_distrib <- stuff[[1]]
      bootstrap_Bca_CI <- stuff[[2]]
    }else
    {
      bootstrap_distrib<-bind_rows(bootstrap_distrib, stuff[[1]])
      bootstrap_Bca_CI<-bind_rows(bootstrap_Bca_CI, stuff[[2]])
    }
  } 
  save(bootstrap_distrib,file=output_file_distrib) 
  save(bootstrap_Bca_CI,file=output_file_Bca)
}


get_from_file_sTfR = TRUE

if (get_from_file_sTfR){
  load(file = output_file_distrib_sTfR)
  load(file = output_file_Bca_sTfR)
}else
{
   for(group_str in c("Pre_menopause_women", "Post_menopause_women","Men")){
    
    stuff <- get_bootstrap_coeffs(blood_data_summary_reg, group_str, outcome = "sTfR", nb_boot = 10000)
    
    if(!exists("bootstrap_distrib_sTfR")){
      bootstrap_distrib_sTfR <- stuff[[1]]
      bootstrap_Bca_CI_sTfR <- stuff[[2]]
    }else
    {
      bootstrap_distrib_sTfR <- bind_rows(bootstrap_distrib_sTfR, stuff[[1]])
      bootstrap_Bca_CI_sTfR <- bind_rows(bootstrap_Bca_CI_sTfR, stuff[[2]])
    }
  } 
  save(bootstrap_distrib_sTfR,file=output_file_distrib_sTfR) 
  save(bootstrap_Bca_CI_sTfR,file=output_file_Bca_sTfR)
}


```

```{r}

# Here we create lists with 95%CI values to include in the final regression model tables.

regressor_values <- c(
 "don_ct" = "Nb donations (2 years)",
   "don_ct_2" = "(Nb donations (2 years))²",
   "log_last_don" = "Days since last donation",
   "iron" = "Iron supplementation",
   "log_CRP" ="CRP (log)",
   "smoking" ="Smoking",
   "age" ="Age",
    "pregnancy" = "Pregnancy",
    "beer_n" = "Beer",
    "red_meat_n" = "Red meat",
    "vegetables_n" = "Vegetables",
    "fruit_berries_n" = "Fruit and Berries",
    "fruit_juices_n" =  "Fruit Juices",
    "milk_n" = "Milk",
    "coffee_n" = "Coffee",
    "tea_n" = "Tea",
    "wine_n" = "Wine", 
    "liquor_n" = "Liquor"
)

## create list for stargazer table

CI_regression_Ferr = list(3)
v_group = c( "Pre_menopause_women", "Post_menopause_women", "Men")

for(i_group in c(1:length(v_group)) ){
  
  CI_matrix <- as.matrix(filter(bootstrap_Bca_CI, group == v_group[i_group]) %>%
                           dplyr::select(Bca_inf, Bca_sup))

  dimnames(CI_matrix) = list(filter(bootstrap_Bca_CI, group == v_group[i_group])$regressor,
    dimnames(CI_matrix)[[2]]) 
  
  CI_regression_Ferr[[i_group]] = CI_matrix
}



bootstrap_Bca_CI <- bootstrap_Bca_CI %>% 
  full_join(lm_signs_std, by=c("regressor", "group")) %>% 
  mutate(regressor = ifelse(regressor == "smokingyes", "smoking", as.character(regressor))) %>% 
  mutate(regressor = ifelse(regressor == "smoking.L", "smoking",as.character(regressor))) %>% 
  mutate(regressor = ifelse(regressor == "pregnancy.L", "pregnancy",as.character(regressor))) %>% 
  mutate(regressor = ifelse(regressor == "pregnancyyes", "pregnancy",as.character(regressor))) %>% 
  mutate(regressor = plyr::revalue(regressor, regressor_values), 
        regressor = ordered(regressor,
                  levels =  c("Age", "BMI", "CRP (log)","Smoking",  "Pregnancy", "Nb donations (2 years)", 
                              "(Nb donations (2 years))²",  "Days since last donation", "Iron supplementation",
                              "Red meat","Vegetables", "Fruit and Berries", "Milk",
                               "Fruit Juices", "Coffee", "Tea", "Beer", "Wine" , "Liquor")),
        regressor = fct_rev(regressor),
        group = case_when(
          group == "Pre_menopause_women" ~ "Pre-menopausal women",
          group == "Post_menopause_women" ~ "Post-menopausal women",
          TRUE ~ "Men"),
        group = ordered(group, levels = c( "Pre-menopausal women", "Post-menopausal women", "Men")))

```
 
```{r}

## create list for stargazer table

CI_regression_sTfR = list(3)
v_group = c( "Pre_menopause_women", "Post_menopause_women", "Men")

for(i_group in c(1:length(v_group)) ){
  
  CI_matrix <- as.matrix(filter(bootstrap_Bca_CI_sTfR, group == v_group[i_group]) %>%
                           dplyr::select(Bca_inf, Bca_sup))

  dimnames(CI_matrix) = list(filter(bootstrap_Bca_CI_sTfR, group == v_group[i_group])$regressor,
    dimnames(CI_matrix)[[2]]) 
  
  CI_regression_sTfR[[i_group]] = CI_matrix
}



bootstrap_Bca_CI_sTfR <- bootstrap_Bca_CI_sTfR %>% 
  full_join(lm_signs_sTfR, by=c("regressor", "group")) %>% 
  mutate(regressor = ifelse(regressor == "smokingyes", "smoking", as.character(regressor))) %>% 
  mutate(regressor = ifelse(regressor == "smoking.L", "smoking",as.character(regressor))) %>% 
  mutate(regressor = ifelse(regressor == "pregnancy.L", "pregnancy",as.character(regressor))) %>% 
  mutate(regressor = ifelse(regressor == "pregnancyyes", "pregnancy",as.character(regressor))) %>% 
  mutate(regressor = plyr::revalue(regressor, regressor_values), 
        regressor = ordered(regressor,
                  levels =  c("Age", "BMI", "CRP (log)","Smoking",  "Pregnancy", "Nb donations (2 years)", 
                              "(Nb donations (2 years))²",  "Days since last donation", "Iron supplementation",
                              "Red meat","Vegetables", "Fruit and Berries", "Milk",
                               "Fruit Juices", "Coffee", "Tea", "Beer", "Wine" , "Liquor")),
        regressor = fct_rev(regressor),
        group = case_when(
          group == "Pre_menopause_women" ~ "Pre-menopausal women",
          group == "Post_menopause_women" ~ "Post-menopausal women",
          TRUE ~ "Men"),
        group = ordered(group, levels = c( "Pre-menopausal women", "Post-menopausal women", "Men")))
```


### Robust regression - coeffs

```{r}
lm7_pre_menop_df_rob <- mdl_pre_menop_w %>% 
  mutate(beta_sign = sign(estimate),
         coefficient = estimate,
   group  = "Pre_menopause_women",
   is_sig = ifelse(p_value < .05, TRUE, FALSE)) %>% 
  dplyr::select(term, beta_sign, group,is_sig,coefficient)

lm7_post_menop_df_rob <- mdl_post_menop_w %>% 
  mutate(beta_sign = sign(estimate),
   group  = "Post_menopause_women", 
            coefficient = estimate,
   is_sig = ifelse(p_value < .05, TRUE, FALSE)) %>%  
  dplyr::select(term, beta_sign, group,is_sig,coefficient)

lm7_men_df_rob <- mdl_men %>% 
  mutate(beta_sign = sign(estimate),
   group  = "Men",
    coefficient = estimate,
   is_sig = ifelse(p_value < .05, TRUE, FALSE)) %>%  
  dplyr::select(term, beta_sign, group,is_sig,coefficient)

lm_signs_rob <- bind_rows(lm7_pre_menop_df_rob,lm7_post_menop_df_rob,lm7_men_df_rob) %>% 
  rename(regressor = term) 

lm7_pre_menop_sTfR_rob <- mdl_sTfR_pre_menop_w %>% 
  mutate(beta_sign = sign(estimate),
         coefficient = estimate,
   group  = "Pre_menopause_women",
   is_sig = ifelse(p_value < .05, TRUE, FALSE)) %>% 
  dplyr::select(term, beta_sign, group,is_sig,coefficient)

lm7_post_menop_sTfR_rob <- mdl_sTfR_post_menop_w %>% 
  mutate(beta_sign = sign(estimate),
   group  = "Post_menopause_women", 
            coefficient = estimate,
   is_sig = ifelse(p_value < .05, TRUE, FALSE)) %>%  
  dplyr::select(term, beta_sign, group,is_sig,coefficient)

lm7_men_sTfR_rob <- mdl_men %>% 
  mutate(beta_sign = sign(estimate),
   group  = "Men",
    coefficient = estimate,
   is_sig = ifelse(p_value < .05, TRUE, FALSE)) %>%  
  dplyr::select(term, beta_sign, group,is_sig,coefficient)

lm_signs_sTfR_rob <- bind_rows(lm7_pre_menop_sTfR_rob,
                               lm7_post_menop_sTfR_rob,
                               lm7_men_sTfR_rob) %>% 
  rename(regressor = term) 

```


```{r }

get_coefficients <- function(data, boot_ind){
  
  model <- tidy(rlm(log_ferr ~ ., data[boot_ind,], maxit = 40)) 
  # %>% filter(term != "(Intercept)")
  return(model$estimate)
}

get_coefficients_sTfR <- function(data, boot_ind){
  
  model <- tidy(rlm(log_sTfR ~ ., data[boot_ind,], maxit = 40)) 
  # %>% filter(term != "(Intercept)")
  return(model$estimate)
}

compute_Bca_CI <-function(fit_boot){
  Bca_inf = rep(0, length(fit_boot$t0))
  Bca_sup = rep(0, length(fit_boot$t0))
  for (i_regressor in 1:length(fit_boot$t0)){
    CI <- boot.ci(fit_boot, type = "bca", index=i_regressor)
    Bca_inf[i_regressor] <- CI$bca[4]
    Bca_sup[i_regressor] <- CI$bca[5]
  }
  
  return(tibble(Bca_inf,Bca_sup)) 
}


get_bootstrap <- function(data, outcome, nb_boot){
  
  if (outcome == "Ferritin"){
      data <- data %>% dplyr::select(-log_sTfR)
      fit_boot <- boot(data, strata = data$smoking, statistic = get_coefficients, R = nb_boot)
  
      fit_boot_distrib <- as.tibble(fit_boot$t)  
      names(fit_boot_distrib) <- tidy(rlm(log_ferr ~ ., data, maxit = 40))$term
  }else{
      data <- data %>% dplyr::select(-log_ferr)
      fit_boot <- boot(data, strata = data$smoking, statistic = get_coefficients_sTfR, R = nb_boot)
  
      fit_boot_distrib <- as.tibble(fit_boot$t)  
      names(fit_boot_distrib) <- tidy(rlm(log_sTfR ~ ., data, maxit = 40))$term
  }

  fit_boot_Bca <- compute_Bca_CI(fit_boot)
  fit_boot_Bca$regressor <- names(fit_boot_distrib)
  
  return(list(fit_boot_distrib,fit_boot_Bca))
}




get_bootstrap_coeffs <- function(blood_data_summary_reg, current_group, outcome, nb_boot )
{
  ## preprocess and standardize data
  test_data_std <- blood_data_summary_reg %>% 
    filter(group == current_group) %>% 
    dplyr::select(donor ,red_meat_n, fruit_berries_n, vegetables_n, milk_n, fruit_juices_n, tea_n, 
                  coffee_n, beer_n, wine_n, liquor_n ) %>% 
    gather(key = key, value = value, -donor) %>% 
    group_by(key) %>% 
    mutate(value = scale(value, scale = FALSE)[,1]) %>% 
    spread(key = key, value = value)
  
  test_data_std <- blood_data_summary_reg %>% 
    filter(group == current_group) %>% 
    dplyr::select(donor, log_ferr, log_sTfR, age, log_CRP, BMI, don_ct,log_last_don,smoking, 
                  pregnancy, iron) %>% 
  mutate(age = age / 5, 
         age = scale(age, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         don_ct = scale(don_ct, scale = FALSE)[,1],
         log_last_don = scale(log_last_don, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         iron = scale(iron, scale = FALSE)[,1],
         don_ct_2 = don_ct^2) %>% 
    full_join(test_data_std, by = "donor")   %>% 
    dplyr::select(-donor) 
  
  if(current_group == "Men"){
    test_data_std <-  dplyr::select(test_data_std, -pregnancy)
  }

  result <- test_data_std %>% 
  get_bootstrap(outcome =  outcome, nb_boot = nb_boot)

  bootstrap_distrib <- result[[1]] %>% 
    gather(key = regressor, value = coefficient) %>% 
    mutate(group = current_group)
    
  Bca_CI  <- result[[2]] %>% mutate(group = current_group)

  return(list(bootstrap_distrib,Bca_CI))
}

set.seed(125)

output_file_distrib = ("../results/bootstraps/coeff_robust_boot_distrib_seed_125_std_strata_final.rdata")
output_file_Bca = ("../results/bootstraps/coeff_robust_boot_Bca_seed_125_std_strata_final.rdata")

output_file_distrib_sTfR = ("../results/bootstraps/coeff_robust_boot_distrib_seed_125_sTfR_final.rdata")
output_file_Bca_sTfR = ("../results/bootstraps/coeff_robust_boot_Bca_seed_125_sTfR_final.rdata")

get_from_file_Ferr = TRUE

if (get_from_file_Ferr){
  load(file=output_file_distrib)
  load(file=output_file_Bca)
}else
{
   for(group_str in c("Pre_menopause_women", "Post_menopause_women","Men")){
    
    stuff <- get_bootstrap_coeffs(blood_data_summary_reg, group_str, outcome =  "Ferritin",  nb_boot = 10000)
    
    if(!exists("bootstrap_distrib_robust")){
      bootstrap_distrib_robust <- stuff[[1]]
      bootstrap_Bca_CI_robust <- stuff[[2]]
    }else
    {
      bootstrap_distrib_robust<-bind_rows(bootstrap_distrib_robust, stuff[[1]])
      bootstrap_Bca_CI_robust<-bind_rows(bootstrap_Bca_CI_robust, stuff[[2]])
    }
  } 
  save(bootstrap_distrib_robust,file=output_file_distrib) 
  save(bootstrap_Bca_CI_robust,file=output_file_Bca)
}

get_from_file_sTfR = TRUE

if (get_from_file_sTfR){
  load(file=output_file_distrib_sTfR)
  load(file=output_file_Bca_sTfR)
}else
{
   for(group_str in c("Pre_menopause_women", "Post_menopause_women","Men")){
    
    stuff <- get_bootstrap_coeffs(blood_data_summary_reg, group_str, outcome =  "sTfR",  nb_boot = 10000)
    
    if(!exists("bootstrap_distrib_robust_sTfR")){
      bootstrap_distrib_robust_sTfR <- stuff[[1]]
      bootstrap_Bca_CI_robust_sTfR <- stuff[[2]]
    }else
    {
      bootstrap_distrib_robust_sTfR<-bind_rows(bootstrap_distrib_robust_sTfR, stuff[[1]])
      bootstrap_Bca_CI_robust_sTfR<-bind_rows(bootstrap_Bca_CI_robust_sTfR, stuff[[2]])
    }
  } 
  save(bootstrap_distrib_robust_sTfR,file=output_file_distrib_sTfR) 
  save(bootstrap_Bca_CI_robust_sTfR,file=output_file_Bca_sTfR)
}
```

```{r}

## Creqate list to include these CIs to the regression results

CI_regression_robust = list(3)
v_group = c( "Pre_menopause_women", "Post_menopause_women", "Men")


# Change the name of smoking and prenancy to match that of robust regression object for 
# proper table output by stargazer later

temp_boot_CI <- 
  bootstrap_Bca_CI_robust %>% 
    mutate(regressor = ifelse(regressor == "smoking", "smoking.L",regressor)) %>%
    mutate(regressor = ifelse(regressor == "pregnancy", "pregnancy.L",regressor)) 

for(i_group in c(1:length(v_group)) ){
  
  CI_matrix <- as.matrix(filter(temp_boot_CI, group == v_group[i_group]) %>% 
                           dplyr::select(Bca_inf, Bca_sup))

  dimnames(CI_matrix) = list(filter(temp_boot_CI, group == v_group[i_group])$regressor,
    dimnames(CI_matrix)[[2]]) 
  
  CI_regression_robust[[i_group]] = CI_matrix
}

bootstrap_Bca_CI_robust <- bootstrap_Bca_CI_robust %>% 
  mutate(regressor = ifelse(regressor == "smoking", "smoking.L",regressor)) %>%
  mutate(regressor = ifelse(regressor == "pregnancy", "pregnancy.L",regressor)) %>%
  full_join(lm_signs_rob, by=c("regressor", "group")) %>% 
  mutate(regressor = ifelse(regressor == "smoking.L", "smoking",regressor)) %>%
  mutate(regressor = ifelse(regressor == "pregnancy.L", "pregnancy",regressor)) %>%
  mutate(regressor = plyr::revalue(regressor, regressor_values), 
        regressor = ordered(regressor,
                  levels =  c("Age", "BMI", "CRP (log)","Smoking",  "Pregnancy", "Nb donations (2 years)", 
                              "(Nb donations (2 years))²",  "Days since last donation", "Iron supplementation",
                              "Red meat","Vegetables", "Fruit and Berries", "Milk",
                               "Fruit Juices", "Coffee", "Tea", "Beer", "Wine" , "Liquor")),
        regressor = fct_rev(regressor),
        group = case_when(
          group == "Pre_menopause_women" ~ "Pre-menopausal women",
          group == "Post_menopause_women" ~ "Post-menopausal women",
          TRUE ~ "Men"),
        group = ordered(group, levels = c( "Men",  "Post-menopausal women", "Pre-menopausal women" )))


```

```{r test1}

## Create list to include these CIs to the regression results

CI_regression_robust_sTfR = list(3)
v_group = c( "Pre_menopause_women", "Post_menopause_women", "Men")


# Change the name of smoking and prenancy to match that of robust regression object for 
# proper table output by stargazer later
temp_boot_CI_sTfR <- 
  bootstrap_Bca_CI_robust_sTfR %>% 
    mutate(regressor = ifelse(regressor == "smoking", "smoking.L",regressor)) %>%
    mutate(regressor = ifelse(regressor == "pregnancy", "pregnancy.L",regressor)) 

for(i_group in c(1:length(v_group)) ){
  
  CI_matrix <- as.matrix(filter(temp_boot_CI_sTfR, group == v_group[i_group]) %>% 
                           dplyr::select(Bca_inf, Bca_sup))

  dimnames(CI_matrix) = list(filter(temp_boot_CI_sTfR, group == v_group[i_group])$regressor,
    dimnames(CI_matrix)[[2]]) 
  
  CI_regression_robust_sTfR[[i_group]] = CI_matrix
}

bootstrap_Bca_CI_robust_sTfR <- bootstrap_Bca_CI_robust_sTfR %>% 
  mutate(regressor = ifelse(regressor == "smoking", "smoking.L",regressor)) %>%
  mutate(regressor = ifelse(regressor == "pregnancy", "pregnancy.L",regressor)) %>%
  full_join(lm_signs_rob, by=c("regressor", "group")) %>% 
  mutate(regressor = ifelse(regressor == "smoking.L", "smoking",regressor)) %>%
  mutate(regressor = ifelse(regressor == "pregnancy.L", "pregnancy",regressor)) %>%
  mutate(regressor = plyr::revalue(regressor, regressor_values), 
        regressor = ordered(regressor,
                  levels =  c("Age", "BMI", "CRP (log)","Smoking",  "Pregnancy", "Nb donations (2 years)", 
                              "(Nb donations (2 years))²",  "Days since last donation", "Iron supplementation",
                              "Red meat","Vegetables", "Fruit and Berries", "Milk",
                               "Fruit Juices", "Coffee", "Tea", "Beer", "Wine" , "Liquor")),
        regressor = fct_rev(regressor),
        group = case_when(
          group == "Pre_menopause_women" ~ "Pre-menopausal women",
          group == "Post_menopause_women" ~ "Post-menopausal women",
          TRUE ~ "Men"),
        group = ordered(group, levels = c( "Men",  "Post-menopausal women", "Pre-menopausal women" )))


```

### Robust regression - betas


```{r test2}

get_coefficients <- function(data, boot_ind){
  
  model <- tidy(rlm(log_ferr ~ ., data[boot_ind,], maxit = 40)) %>% filter(term != "(Intercept)")
  return(model$estimate)
}

get_coefficients_sTfR <- function(data, boot_ind){
  
  model <- tidy(rlm(log_sTfR ~ ., data[boot_ind,], maxit = 40)) %>% filter(term != "(Intercept)")
  return(model$estimate)
}

compute_Bca_CI <-function(fit_boot, outcome){
  Bca_inf = rep(0, length(fit_boot$t0))
  Bca_sup = rep(0, length(fit_boot$t0))
  for (i_regressor in 1:length(fit_boot$t0)){
    CI <- boot.ci(fit_boot, type = "bca", index=i_regressor)
    Bca_inf[i_regressor] <- CI$bca[4]
    Bca_sup[i_regressor] <- CI$bca[5]
  }
  
  if (outcome == "Ferritin"){
   regressor <- fit_boot$data %>% dplyr::select(-log_ferr) %>% names()
  }else{
   regressor <- fit_boot$data %>% dplyr::select(-log_sTfR) %>% names()
  }
  
  return(tibble(Bca_inf,Bca_sup, regressor = regressor)) 
         
}


get_bootstrap <- function(data, outcome, nb_boot){
  
   if (outcome == "Ferritin"){
     
    data <- data %>% dplyr::select(-log_sTfR)
     
    fit_boot <- boot(data, strata = data$smoking, statistic= get_coefficients, R = nb_boot)
  
    fit_boot_distrib <- as.tibble(fit_boot$t)  
    names(fit_boot_distrib) <- fit_boot$data %>% dplyr::select(-log_ferr) %>% names()
    
   }else{
     
    data <- data %>% dplyr::select(-log_ferr)
     
    fit_boot <- boot(data, strata = data$smoking, statistic= get_coefficients_sTfR, R = nb_boot)
  
    fit_boot_distrib <- as.tibble(fit_boot$t)  
    names(fit_boot_distrib) <- fit_boot$data %>% dplyr::select(-log_sTfR) %>% names()
    
   }


  fit_boot_Bca <- compute_Bca_CI(fit_boot, outcome = outcome)
  
  return(list(fit_boot_distrib,fit_boot_Bca))
}

preprocess_data <- function(blood_data_summary_reg, current_group){
  
    ## preprocess and standardize data
  test_data_std <- blood_data_summary_reg %>% 
    filter(group == current_group) %>% 
    dplyr::select(donor ,red_meat_n, fruit_berries_n, vegetables_n, milk_n, fruit_juices_n, tea_n, 
                  coffee_n, beer_n, wine_n, liquor_n ) %>% 
    gather(key = key, value = value, -donor) %>% 
    group_by(key) %>% 
    mutate(value = scale(value, scale = TRUE)[,1]) %>% 
    spread(key = key, value = value)
  
  test_data_std <- blood_data_summary_reg %>% 
    filter(group == current_group) %>% 
    dplyr::select(donor, log_ferr, log_sTfR, age, log_CRP, BMI, don_ct, log_last_don, smoking, 
                  pregnancy, iron) %>% 
  mutate(age = age / 5, 
         age = scale(age, scale = TRUE)[,1],
         log_CRP = scale(log_CRP, scale = TRUE)[,1],
         don_ct = scale(don_ct, scale = TRUE)[,1],
         log_last_don = scale(log_last_don, scale = TRUE)[,1],
         BMI = scale(BMI, scale = TRUE)[,1],
         iron = scale(iron, scale = TRUE)[,1],
         log_ferr = scale(log_ferr, scale = TRUE)[,1],
         log_sTfR = scale(log_sTfR, scale = TRUE)[,1],
         don_ct_2 = don_ct^2) %>% 
    full_join(test_data_std, by = "donor")   %>% 
    dplyr::select(-donor) 
  
  if(current_group == "Men"){
    test_data_std <-  dplyr::select(test_data_std, -pregnancy)
  }

  return(test_data_std)
}


get_bootstrap_coeffs <- function(test_data_std, current_group, outcome, nb_boot )
{
  result <- test_data_std %>% 
  get_bootstrap(outcome = outcome, nb_boot = nb_boot)

  bootstrap_distrib <- result[[1]] %>% 
    gather(key = regressor, value = coefficient) %>% 
    filter(regressor != "(Intercept)") %>% 
    mutate(group = current_group)
    
  Bca_CI  <- result[[2]] %>% mutate(group = current_group)

  return(list(bootstrap_distrib,Bca_CI))
}

set.seed(3547)

output_file_distrib = ("../results/bootstraps/beta_robust_boot_distrib_seed_3547_std_strata.rdata")
output_file_Bca = ("../results/bootstraps/beta_robust_boot_Bca_seed_3547_std_strata.rdata")
# 
output_file_distrib_sTfR = ("../results/bootstraps/beta_robust_boot_distrib_seed_3547_stfR_strata.rdata")
output_file_Bca_sTfR = ("../results/bootstraps/beta_robust_boot_Bca_seed_3547_stfR_strata.rdata")

get_from_file_Ferr = TRUE

if (get_from_file_Ferr){
  load(file=output_file_distrib)
  load(file=output_file_Bca)
}else
{
   for(group_str in c("Pre_menopause_women", "Post_menopause_women","Men")){
    
    test_data_std <- preprocess_data(blood_data_summary_reg, group_str)
    stuff <- get_bootstrap_coeffs(test_data_std, group_str, outcome = "Ferritin", nb_boot = 10000)
    
    #Remove sTfR from the data object because of model estimation
    test_data_temp <- test_data_std %>% dplyr::select(-log_sTfR)
    
    bootstrap_Bca_CI_temp <- tidy(rlm(log_ferr ~ ., test_data_temp, maxit = 40)) %>%
      filter(term != "(Intercept)") %>%
      mutate(term = ifelse(term == "smoking.L","smoking",term),
             term = ifelse(term == "pregnancy.L","pregnancy",term))%>% 
      dplyr::select(term, estimate) %>% 
      rename(regressor = term, coefficient = estimate) %>% 
      full_join(stuff[[2]], by = "regressor")
      
      
    if(!exists("bootstrap_distrib_robust_std")){
      bootstrap_distrib_robust_std <- stuff[[1]]
      bootstrap_Bca_CI_robust_std <- bootstrap_Bca_CI_temp
    }else
    {
      bootstrap_distrib_robust_std <- bind_rows(bootstrap_distrib_robust_std, stuff[[1]])
      bootstrap_Bca_CI_robust_std <- bind_rows(bootstrap_Bca_CI_robust_std, bootstrap_Bca_CI_temp)
    }
  } 
  save(bootstrap_distrib_robust_std,file=output_file_distrib) 
  save(bootstrap_Bca_CI_robust_std,file=output_file_Bca)
}

bootstrap_Bca_CI_robust_std <-
  bootstrap_Bca_CI_robust_std %>% 
  mutate(regressor = plyr::revalue(regressor, regressor_values),
        regressor = ordered(regressor,
                  levels =  c("Age","CRP (log)", "BMI","Nb donations (2 years)","(Nb donations (2 years))²", 
                              "Days since last donation",  "Smoking",  "Pregnancy", "Iron supplementation", 
                              "Red meat","Vegetables", "Fruit and Berries", "Milk",
                         "Fruit Juices", "Coffee", "Tea", "Beer", "Wine" , "Liquor")),
        regressor = fct_rev(regressor),
        group = case_when(group == "Pre_menopause_women" ~ "Pre-menopausal women",
                          group == "Post_menopause_women" ~ "Post-menopausal women",
                          TRUE ~ "Men"),
        group = ordered(group, levels = c( "Men",  "Post-menopausal women", "Pre-menopausal women" )))


get_from_file_sTfR = TRUE

if (get_from_file_sTfR){
  load(file=output_file_distrib_sTfR)
  load(file=output_file_Bca_sTfR)
}else
{
   for(group_str in c("Pre_menopause_women", "Post_menopause_women","Men")){
    
    test_data_std <- preprocess_data(blood_data_summary_reg, group_str)
    stuff <- get_bootstrap_coeffs(test_data_std, group_str, outcome = "sTfR", nb_boot = 10000)

       #Remove Ferritin from the data object because of model estimation
    test_data_temp <- test_data_std %>% dplyr::select(-log_ferr)

        
    bootstrap_Bca_CI_temp <- tidy(rlm(log_sTfR ~ ., test_data_temp, maxit = 40)) %>%
      filter(term != "(Intercept)") %>%
      mutate(term = ifelse(term == "smoking.L","smoking",term),
             term = ifelse(term == "pregnancy.L","pregnancy",term))%>% 
      dplyr::select(term, estimate) %>% 
      rename(regressor = term, coefficient = estimate) %>% 
      full_join(stuff[[2]], by = "regressor")
      
      
    if(!exists("bootstrap_distrib_robust_std_sTfR")){
      bootstrap_distrib_robust_std_sTfR <- stuff[[1]]
      bootstrap_Bca_CI_robust_std_sTfR <- bootstrap_Bca_CI_temp
    }else
    {
      bootstrap_distrib_robust_std_sTfR <- bind_rows(bootstrap_distrib_robust_std_sTfR, stuff[[1]])
      bootstrap_Bca_CI_robust_std_sTfR <- bind_rows(bootstrap_Bca_CI_robust_std_sTfR, bootstrap_Bca_CI_temp)
    }
  } 
  save(bootstrap_distrib_robust_std_sTfR,file=output_file_distrib_sTfR) 
  save(bootstrap_Bca_CI_robust_std_sTfR,file=output_file_Bca_sTfR)
}

bootstrap_Bca_CI_robust_std_sTfR <- bootstrap_Bca_CI_robust_std_sTfR %>% 
  mutate(regressor = plyr::revalue(regressor, regressor_values),
        regressor = ordered(regressor,
                  levels =  c("Age","CRP (log)", "BMI","Nb donations (2 years)","(Nb donations (2 years))²", 
                              "Days since last donation",  "Smoking",  "Pregnancy", "Iron supplementation", 
                              "Red meat","Vegetables", "Fruit and Berries", "Milk",
                         "Fruit Juices", "Coffee", "Tea", "Beer", "Wine" , "Liquor")),
        regressor = fct_rev(regressor),
        group = case_when(group == "Pre_menopause_women" ~ "Pre-menopausal women",
                          group == "Post_menopause_women" ~ "Post-menopausal women",
                          TRUE ~ "Men"),
        group = ordered(group, levels = c( "Men",  "Post-menopausal women", "Pre-menopausal women" )))
```

## Results outputs

### Main results 

#### Table 2 - Outcome = Ferritin 

```{r}
# create pvlue list to get the p values from the appropriate package

robust_p_values = list(3)

p_matrix <- as.matrix(mdl_pre_menop_w %>% 
                           dplyr::select(p_value))
dimnames(p_matrix) = list(mdl_pre_menop_w$term) 
  robust_p_values[[1]] = p_matrix

p_matrix <- as.matrix(mdl_post_menop_w %>% 
                           dplyr::select(p_value))
dimnames(p_matrix) = list(mdl_post_menop_w$term) 
robust_p_values[[2]] = p_matrix

p_matrix <- as.matrix(mdl_men %>% 
                           dplyr::select(p_value))
dimnames(p_matrix) = list(mdl_men$term) 
robust_p_values[[3]] = p_matrix

```


```{r, results='asis'}

# ouput to file

stargazer(robust_mdl_pre_menop_w, robust_mdl_post_menop_w, robust_mdl_men,
          out = "../results/tables/table_2_robust_regressions_complete__values.html",
          style = "all", 
          intercept.bottom = FALSE,
          title = "Table 2 - Multivariate robust regression analyses - Ferritin",
          type = "html",
          # se = TRUE,
          single.row = TRUE,
          ci = TRUE,
          ci.custom = CI_regression_robust,
          p = robust_p_values,
          omit = "Constant",
          header = FALSE,
          covariate.labels = c("Age", "BMI", "ln(CRP)", "Smoking(yes)", "Pregnancy(Yes)",
                       "Nb donations (2 years)","(Nb donations (2 years))²", "ln(Days since last donation)",
                        "iron supplementation", "Red meat","Vegetables", "Fruit and Berries", "Milk",
                         "Fruit Juices", "Coffee","Tea", "Beer", "Wine" , "Liquor"),
          dep.var.caption   = "Ferritin (log)", 
          dep.var.labels.include = FALSE,
          column.labels = c("Pre-menopausal women", "Post-menopausal women", "Men"),
          model.numbers= FALSE,
          report=('vcsp'),
          digits = 3)
```

- 

#### Table 3 - Outcome = sTfR


```{r}
# create pvlue list to get the p values from the appropriate package

robust_p_values_sTfR = list(3)

p_matrix <- as.matrix(mdl_sTfR_pre_menop_w %>% 
                           dplyr::select(p_value))
dimnames(p_matrix) = list(mdl_sTfR_pre_menop_w$term) 
  robust_p_values_sTfR[[1]] = p_matrix

p_matrix <- as.matrix(mdl_sTfR_post_menop_w %>% 
                           dplyr::select(p_value))
dimnames(p_matrix) = list(mdl_sTfR_post_menop_w$term) 
robust_p_values_sTfR[[2]] = p_matrix

p_matrix <- as.matrix(mdl_sTfR_men %>% 
                           dplyr::select(p_value))
dimnames(p_matrix) = list(mdl_sTfR_men$term) 
robust_p_values_sTfR[[3]] = p_matrix

```


```{r, results='asis'}

stargazer(robust_mdl_sTfR_pre_menop_w, robust_mdl_sTfR_post_menop_w, robust_mdl_sTfR_men,
          out = "../results/tables/table_3_robust_regressions_complete__values.html",
          style = "all", 
          intercept.bottom = FALSE,
          title = "Table 3 - Multivariable robust regression analyses - sTfR",
          type = "html",
          # se = TRUE,
          single.row = TRUE,
          ci = TRUE,
          ci.custom = CI_regression_robust_sTfR,
          p = robust_p_values,
          omit = "Constant",
          header = FALSE,
          covariate.labels = c("Age", "BMI", "ln(CRP)", "Smoking(yes)", "Pregnancy(Yes)",
                       "Nb donations (2 years)","(Nb donations (2 years))²", "ln(Days since last donation)",
                        "iron supplementation", "Red meat","Vegetables", "Fruit and Berries", "Milk",
                         "Fruit Juices", "Coffee","Tea", "Beer", "Wine" , "Liquor"),
          dep.var.caption   = "sTfR (log)", 
          dep.var.labels.include = FALSE,
          column.labels = c("Pre-menopausal women", "Post-menopausal women", "Men"),
          model.numbers= FALSE,
          report=('vcsp'),
          digits = 3)
```

-

#### Figure 2 

##### Bootstrapped coeffs

The bootstrapped coefficients that are displayed are the standardized coefficients. 

###### Ferritin

```{r}

bootstrap_distrib_robust_std %>% 
  mutate(regressor = plyr::revalue(regressor, regressor_values),
  regressor = ordered(regressor,
                  levels =  c("Age", "BMI", "CRP (log)",  "Smoking",  "Pregnancy","Nb donations (2 years)",
                              "(Nb donations (2 years))²", "Days since last donation",  "Iron supplementation", 
                              "Red meat","Vegetables", "Fruit and Berries", "Milk",
                         "Fruit Juices", "Coffee", "Tea", "Beer", "Wine" , "Liquor")),
  regressor = fct_rev(regressor),
  group = case_when(group == "Pre_menopause_women" ~ "Pre-menopausal women",
    group == "Post_menopause_women" ~ "Post-menopausal women",
    TRUE ~ "Men"),
  group = ordered(group, levels = c( "Men",  "Post-menopausal women", "Pre-menopausal women" ) )) %>% 
  ggplot(aes(x = regressor, y = coefficient, color = group, fill = group)) +
  geom_hline(yintercept = 0 , color = "grey", linetype = "dotted") +
  geom_violin(alpha = 0.15, trim = TRUE, position = position_dodge(width = 0.8), size = 0.1) +
  geom_linerange(data = bootstrap_Bca_CI_robust_std, aes(ymin=Bca_inf, ymax = Bca_sup ),
                  position = position_dodge(width = 0.8), size = 0.4) +
  geom_point(data = bootstrap_Bca_CI_robust_std, position = position_dodge(width = 0.8), size = 0.75, shape = 4) +
  scale_color_manual(values = c( "#00BFFF",  "#de2d26", "#ff85a2" ),
                     limits = c( "Men",  "Post-menopausal women", "Pre-menopausal women" )) +
  scale_fill_manual(values = c( "#00BFFF",  "#de2d26", "#ff85a2" ),
                     limits = c( "Men",  "Post-menopausal women", "Pre-menopausal women" )) +
  theme_classic() +
  theme(legend.position = "bottom",
        panel.grid.minor.y = element_blank(),
        axis.line = element_line(colour="black"),
        axis.title.y = element_blank(),
        axis.title.x = element_text(size = 8),
        axis.text = element_text(size = 6),
        panel.grid.major.y = element_blank()) +
  coord_flip() +
  xlab("Robust standardized coefficient")

ggplot2::ggsave("../results/figures/Figure_2_coeffs_Ferritin.png",device = "png", height = 17, width = 6.75, units = "cm" )

```


###### sTfR


```{r}

test <- bootstrap_distrib_robust_std_sTfR %>% 
  mutate(regressor = plyr::revalue(regressor, regressor_values),
  regressor = ordered(regressor,
                  levels =  c("Age", "BMI", "CRP (log)",  "Smoking",  "Pregnancy","Nb donations (2 years)",
                              "(Nb donations (2 years))²", 
                              "Days since last donation",  "Iron supplementation", 
                              "Red meat","Vegetables", "Fruit and Berries", "Milk",
                         "Fruit Juices", "Coffee", "Tea", "Beer", "Wine" , "Liquor")),
  regressor = fct_rev(regressor))

```

```{r}

bootstrap_distrib_robust_std_sTfR %>% 
  mutate(regressor = plyr::revalue(regressor, regressor_values),
  regressor = ordered(regressor,
                  levels =  c("Age", "BMI", "CRP (log)",  "Smoking",  "Pregnancy","Nb donations (2 years)",
                              "(Nb donations (2 years))²", 
                              "Days since last donation",  "Iron supplementation", 
                              "Red meat","Vegetables", "Fruit and Berries", "Milk",
                         "Fruit Juices", "Coffee", "Tea", "Beer", "Wine" , "Liquor")),
  regressor = fct_rev(regressor),
  group = case_when(group == "Pre_menopause_women" ~ "Pre-menopausal women",
    group == "Post_menopause_women" ~ "Post-menopausal women",
    TRUE ~ "Men"),
  group = ordered(group, levels = c( "Men",  "Post-menopausal women", "Pre-menopausal women" ) )) %>% 
  ggplot(aes(x = regressor, y = coefficient, color = group, fill = group)) +
  geom_hline(yintercept = 0 , color = "grey", linetype = "dotted") +
  geom_violin(alpha = 0.15, trim = TRUE, position = position_dodge(width = 0.8), size = 0.1) +
  geom_linerange(data = bootstrap_Bca_CI_robust_std_sTfR, aes(ymin=Bca_inf, ymax = Bca_sup ),
                  position = position_dodge(width = 0.8), size = 0.4) +
  geom_point(data = bootstrap_Bca_CI_robust_std_sTfR, position = position_dodge(width = 0.8), size = 0.75, shape = 4) +
  scale_color_manual(values = c( "#00BFFF",  "#de2d26", "#ff85a2" ),
                     limits = c( "Men",  "Post-menopausal women", "Pre-menopausal women" )) +
  scale_fill_manual(values = c( "#00BFFF",  "#de2d26", "#ff85a2" ),
                     limits = c( "Men",  "Post-menopausal women", "Pre-menopausal women" )) +
  theme_classic() +
  theme(legend.position = "bottom",
        panel.grid.minor.y = element_blank(),
        axis.line = element_line(colour="black"),
        axis.title.y = element_blank(),
        axis.title.x = element_text(size = 8),
        axis.text = element_text(size = 6),
        panel.grid.major.y = element_blank()) +
  coord_flip() +
  xlab("Robust standardized coefficient")

ggplot2::ggsave("../results/figures/Figure_2_coeffs_sTfR.png",device = "png", height = 17, width = 6.75, units = "cm" )

```


##### Relative importance

###### Ferritin

```{r message=FALSE, warning=FALSE}
plot_data_lm7_men$group <- "Men"
plot_data_lm7_pre_menop$group <- "Pre_menopause_women"
plot_data_lm7_post_menop$group <- "Post_menopause_women"

relativ_impo_data <- bind_rows(plot_data_lm7_men,
                              plot_data_lm7_pre_menop,
                             plot_data_lm7_post_menop) %>% 
mutate(group = factor(group, levels =  c("Pre_menopause_women", "Post_menopause_women", "Men")))
  
relativ_impo_data <- relativ_impo_data %>% 
  mutate(regressor = ifelse(regressor == "smoking", "smokingyes", regressor)) %>% 
  full_join(lm_signs_std, by = c("regressor", "group")) %>% 
  mutate(regressor = ifelse(regressor == "smokingyes", "smoking", regressor)) %>%
  dplyr::filter(regressor %in% c("age", "BMI","log_CRP", "don_ct", "don_ct_2", "log_last_don", 
                                "smoking", "iron", "red_meat_n", "milk_n", "beer_n", "wine_n")) %>% 
  mutate(lmg = lmg *beta_sign * 100,
         lmg_inf = lmg_inf * beta_sign * 100,
         lmg_sup = lmg_sup * beta_sign * 100)

dodge <- position_dodge(width=0.9)

plt <- relativ_impo_data %>% 
  mutate(regressor = case_when(
  regressor == "don_ct" ~ "Nb donations (2 years)",
  regressor == "don_ct_2" ~ "(Nb donations (2 years))²",
  regressor == "log_last_don" ~ "Time since last donation",
  regressor == "iron" ~ "Iron supplementation",
  regressor == "log_CRP" ~"CRP",
  regressor == "smoking" ~"Smoking",
  regressor == "age" ~"Age",
  regressor == "milk_n" ~"Milk",
  regressor == "beer_n" ~"Beer",
  regressor == "wine_n" ~"Wine",
  regressor == "red_meat_n" ~"Red meat",
  TRUE ~regressor)) %>% 
  mutate(regressor = ordered(regressor, 
                  levels = c("Wine", "Beer", "Red meat", "Milk", "Iron supplementation",  
                             "Time since last donation",  "(Nb donations (2 years))²",
                             "Nb donations (2 years)","Smoking", "CRP", "BMI",  "Age"))) %>%
  mutate(group_2 = case_when(
    group == "Pre_menopause_women" ~ "Pre-menopausal women",
    group == "Post_menopause_women" ~ "Post-menopausal women",
    TRUE ~ "Men"),
    group_2 = ordered(group_2, levels = c(  "Men", "Post-menopausal women", "Pre-menopausal women") )) %>% 
  ggplot(aes(x=regressor, y = lmg, fill = group_2)) +
  geom_col(position = dodge) +
  geom_linerange(aes( ymin=lmg_inf, ymax = lmg_sup), position = dodge, size = 0.4, color = "dark grey") +
  scale_fill_manual(values = c( "#ff85a2",  "#de2d26","#00BFFF" ),
                     limits = c("Pre-menopausal women", "Post-menopausal women","Men" )) +
  coord_flip() +
  theme_bw() +
   theme(legend.position = "none", 
        panel.grid.minor.y = element_blank(),
        panel.border = element_rect(colour = NA),
        strip.background=element_rect(colour="white",fill="white"),
        strip.text = element_text(face="bold", size = 8*1.333333), # weird sizing issue ...
        axis.line = element_line(colour="black"),
        axis.title.y = element_blank(),
        axis.title.x = element_text(size = 8*1.3333),
        axis.text = element_text(size = 6*1.333333),
        panel.grid.major.y = element_blank()) +
  ylab("Average percentage of explained \nvariance (pmvd)")

plt 

ggplot2::ggsave("../results/figures/figure_2_rela_impo_ferritin.svg", plt,  height = 6.75 * 1.25, 
                width = 6.75 * 1.25, units = "cm", dpi = 300 )
```

###### sTfR

```{r warning=FALSE, message=FALSE}
plot_data_lm7_men_sTfR$group <- "Men"
plot_data_lm7_pre_menop_sTfR$group <- "Pre_menopause_women"
plot_data_lm7_post_menop_sTfR$group <- "Post_menopause_women"

relativ_impo_data_sTfR <- bind_rows(plot_data_lm7_men_sTfR,
                              plot_data_lm7_pre_menop_sTfR,
                             plot_data_lm7_post_menop_sTfR) %>% 
mutate(group = factor(group, levels =  c("Pre_menopause_women", "Post_menopause_women", "Men")))
  
relativ_impo_data_sTfR <- relativ_impo_data_sTfR %>% 
  mutate(regressor = ifelse(regressor == "smoking", "smokingyes", regressor)) %>% 
  full_join(lm_signs_sTfR, by = c("regressor", "group")) %>% 
  mutate(regressor = ifelse(regressor == "smokingyes", "smoking", regressor)) %>%
  dplyr::filter(regressor %in% c("age", "BMI","log_CRP", "don_ct", "don_ct_2", "log_last_don", 
                                "smoking", "iron", "red_meat_n", "milk_n", "beer_n", "wine_n")) %>% 
  mutate(lmg = lmg *beta_sign * 100,
         lmg_inf = lmg_inf * beta_sign * 100,
         lmg_sup = lmg_sup * beta_sign * 100)

dodge <- position_dodge(width=0.9)

plt <- relativ_impo_data_sTfR %>% 
  mutate(regressor = case_when(
  regressor == "don_ct" ~ "Nb donations (2 years)",
  regressor == "don_ct_2" ~ "(Nb donations (2 years))²",
  regressor == "log_last_don" ~ "Time since last donation",
  regressor == "iron" ~ "Iron supplementation",
  regressor == "log_CRP" ~"CRP",
  regressor == "smoking" ~"Smoking",
  regressor == "age" ~"Age",
  regressor == "wine_n" ~"Wine",
  regressor == "red_meat_n" ~ "Red meat",
  regressor == "milk_n" ~ "Milk",
  regressor == "beer_n" ~ "Beer",
  TRUE ~regressor)) %>% 
  mutate(regressor = ordered(regressor, 
                  levels = c("Wine", "Beer", "Red meat", "Milk", "Iron supplementation",  
                             "Time since last donation",  "(Nb donations (2 years))²",
                             "Nb donations (2 years)", "Smoking", "CRP", "BMI",  "Age"))) %>%
  mutate(group_2 = case_when(
    group == "Pre_menopause_women" ~ "Pre-menopausal women",
    group == "Post_menopause_women" ~ "Post-menopausal women",
    TRUE ~ "Men"),
          group_2 = ordered(group_2, levels = c(  "Men", "Post-menopausal women", "Pre-menopausal women") )) %>% 
  ggplot(aes(x=regressor, y = lmg, fill = group_2)) +
  geom_col(position = dodge) +
  geom_linerange(aes( ymin=lmg_inf, ymax = lmg_sup), position = dodge, size = 0.4, color = "dark grey") +
  scale_fill_manual(values = c( "#ff85a2",  "#de2d26","#00BFFF" ),
                     limits = c("Pre-menopausal women", "Post-menopausal women","Men" )) +
  coord_flip() +
  theme_bw() +
   theme(legend.position = "none", 
        panel.grid.minor.y = element_blank(),
        panel.border = element_rect(colour = NA),
        strip.background=element_rect(colour="white",fill="white"),
        strip.text = element_text(face="bold", size = 8*1.333333), # weird sizing issue ...
        axis.line = element_line(colour="black"),
        axis.title.y = element_blank(),
        axis.title.x = element_text(size = 8*1.3333),
        axis.text = element_text(size = 6*1.333333),
        panel.grid.major.y = element_blank()) +
  ylab("Average percentage of explained \nvariance (pmvd)")

plt 

ggplot2::ggsave("../results/figures/figure_2_rela_impo_sTfR.svg", plt,  height = 6.75 * 1.25, width = 6.75 * 1.25,
                units = "cm", dpi = 300 )
```

#### Figure 3 : Effects of main predictors of interest 

```{r , warning=FALSE}
plt_1 <-blood_data_summary_final %>%
  dplyr::select(donor, group2, Ferritin, 
         TwoYearsFromStartCount_FB) %>% 
  ggplot(aes(x=TwoYearsFromStartCount_FB, y=Ferritin)) +
  geom_hline(data = threshold, aes(yintercept = Ferritin_thresh), color = "dark red", size = .2) +
  geom_jitter(aes(color = group2), alpha = 0.25, width = 0.2, size = 0.2, height = 0, shape = 1) +
  geom_boxplot(aes(color = group2, group = TwoYearsFromStartCount_FB), alpha = 0,
               width = 0.4,
               outlier.alpha = 0) +
  facet_grid(group2~.) +
  scale_color_manual(values = c( "#ff85a2",  "#de2d26","#00BFFF" ),
                     limits = c("Pre-menopausal women", "Post-menopausal women","Men" )) +
  theme_bw() +
  theme(legend.position = "none", 
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
       panel.border = element_rect(colour = NA),
        strip.background=element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(colour="black"),
        axis.title.x = element_text(size = 8),
        axis.title.y =element_text(size = 8),
        axis.text = element_text(size = 8)) +
  scale_y_log10(breaks=c(5, 15,25,50,100)) +
  scale_x_continuous(breaks=c(0, 2, 4,6,8, 10, 12)) +
  xlab("Nb of donations (2 years)") +
  ylab(expression("Ferritin ("*mu* "g/l)")) 

plt_2 <-blood_data_summary_final %>%
  dplyr::select(donor, group2, Ferritin, 
         DaysToPreviousFB) %>% 
 ggplot(aes(x=DaysToPreviousFB, y=Ferritin)) +
  geom_hline(data = threshold, aes(yintercept = Ferritin_thresh), color = "dark red", size = .2) +
  geom_point(aes(color = group2), alpha = 0.4,  size = 0.7, shape =1) +
  facet_grid(group2~., 
      labeller = label_wrap_gen(width = 2, multi_line = TRUE))  +
  scale_color_manual(values = c( "#ff85a2",  "#de2d26","#00BFFF" ),
                     limits = c("Pre-menopausal women", "Post-menopausal women","Men" )) +
  theme_bw() +
  theme(legend.position = "none", 
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = NA),
        strip.background=element_rect(colour="white",fill="white"),
        strip.text = element_text(face="bold", size = 10),
        axis.line = element_line(colour="black"),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_blank(),
        axis.text = element_text(size = 8)) +
  scale_y_log10(breaks=c(5, 15,25,50,100)) +
  scale_x_log10(breaks=c(60, 90,180,360,1000)) +
  xlab("Days since last donation") +
  ylab("")

plt_3 <- blood_data_summary_final %>% 
  mutate(iron_comp_f = case_when(
    iron_comp_c == 3 ~ "All or almost all",
    iron_comp_c == 2 ~ "About half",
    iron_comp_c == 1 ~ "Less than half",
    iron_comp_c == 0 ~ "None",
    TRUE ~ "NA"),
  iron_comp_f = ordered(iron_comp_f, 
                  levels = c("NA","None", "Less than half",  "About half","All or almost all" ))) %>% 
  dplyr::select(group2, iron_comp_f, Ferritin) %>% 
  ggplot(aes(x = iron_comp_f, y = Ferritin, color = group2 )) +
  geom_hline(data = threshold, aes(yintercept = Ferritin_thresh), color = "dark red", size = .2) +
  geom_boxplot(outlier.alpha = 0, position = "dodge", width = 0.5) +
  geom_jitter(width = 0.25, height = 0, shape = 1, size = 0.2, alpha = 0.25) +
  scale_color_manual(values = c( "#ff85a2",  "#de2d26","#00BFFF" ),
                     limits = c( "Pre-menopausal women", "Post-menopausal women", "Men" )) +
  facet_grid(group2~., 
      labeller = label_wrap_gen(width = 2, multi_line = TRUE)) +
 theme(legend.position = "none", 
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
       panel.border = element_rect(colour = NA),
        strip.background=element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(colour="black"),
        axis.text.x=element_text(angle=20,hjust=1),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        axis.text = element_text(size = 8)) +
  scale_y_log10(breaks=c(5, 15,25,50,100)) +
  xlab("Iron supplementation")  + 
  ylab(expression("Ferritin ("*mu* "g/l)")) 

plt_4 <- blood_data_summary_final %>% 
    mutate(red_meat = case_when(red_meat == "never" ~  "Never",
    red_meat == "less_than_once_weekly" ~ "Less than weekly",
    red_meat %in% c("1.3_week" , "4.6_week") ~ "Several weekly",
    TRUE   ~ "Daily or more" ), 
    red_meat = ordered(red_meat, 
                  levels = c("Never", "Less than weekly", "Several weekly", "Daily or more" ))) %>% 
  dplyr::select(group2, red_meat, Ferritin) %>% 
  ggplot(aes(x = red_meat, y = Ferritin, color = group2 )) +
  geom_hline(data = threshold, aes(yintercept = Ferritin_thresh), color = "dark red", size = .2) +
  geom_boxplot(outlier.alpha = 0, position = "dodge", width = 0.5) +
  geom_jitter(width = 0.25, height = 0, shape = 1, size = 0.2, alpha = 0.3) +
  scale_color_manual(values = c( "#ff85a2",  "#de2d26","#00BFFF" ),
                     limits = c( "Pre-menopausal women", "Post-menopausal women", "Men" )) +
  facet_grid(group2~., 
      labeller = label_wrap_gen(width = 2, multi_line = TRUE)) +
 theme(legend.position = "none", 
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = NA),
        strip.background=element_rect(colour="white",fill="white"),
        strip.text = element_text(face="bold", size = 10),
        axis.line = element_line(colour="black"),
        axis.text.x=element_text(angle=20,hjust=1),
        axis.title.x =  element_text(size = 8),
       axis.title.y =  element_blank(),
        axis.text = element_text(size = 8)) +
  scale_y_log10(breaks=c(5, 15,25,50,100)) +
  xlab("Red meat") 

plt2 <- cowplot::plot_grid(plt_1, plt_2, plt_3, plt_4,  ncol = 2, labels = c("A", "B", "C", "D"))

plt2

cowplot::save_plot("../results/figures/Figure_3_predictors.png",plt2, base_height = NULL,
          base_aspect_ratio = .5, base_width = 11.5 * .394 ) # make room for figure legend
```


## Supplementary results


### Supp Table 1: OLS ferritin regression

```{r, results='asis'}
stargazer(lm7_pre_menop_std, lm7_post_menop_std, lm7_men_std,
          out = "../results/tables/Supp_table_3_OLS_regressions_with_p.html",
          style = "all", 
          intercept.bottom = FALSE,
          title = "Supplementary table 3 - Multivariate linear regression (OLS) - Ferritin",
          type = "html",
          # se = TRUE,
          single.row = TRUE,
          ci = TRUE,
          ci.custom = CI_regression_Ferr,
          omit = "Constant",
          header = FALSE,
          covariate.labels = c("Age", "BMI", "ln(CRP)", "Smoking(yes)", "Pregnancy(Yes)",
                       "Nb donations (2 years)","(Nb donations (2 years))²", "ln(Days since last donation)",
                        "iron supplementation", "Red meat","Vegetables", "Fruit and Berries", "Milk",
                         "Fruit Juices", "Coffee","Tea", "Beer", "Wine" , "Liquor"),
          dep.var.caption   = "Ferritin (log)", 
          dep.var.labels.include = FALSE,
          column.labels = c("Pre-menopausal women", "Post-menopausal women", "Men"),
          model.numbers= FALSE,
          report=('vcsp'),
          star.cutoffs = c(0.05, 0.01, 0.001),
          digits = 2)
```

-

### Supp Table 2: Regression of sTfR

```{r, results='asis'}
stargazer(lm7_pre_menop_sTfR, lm7_post_menop_sTfR,lm7_men_sTfR,
          out = "../results/tables/supp_table_2_sTfR_regressions_coeffs_OLS.html",
          intercept.bottom = FALSE,
          title = "Supplementary Table 3 - Multivariable linear regression Analyses (sTfR)",
          type = "html",
          single.row = TRUE,
          ci = TRUE,
          ci.custom = CI_regression_sTfR,
          omit = "Constant",
          header = FALSE,
          covariate.labels = c("Age", "BMI", "CRP", "Smoking(yes)", "Pregnancy(Yes)",
                       "Nb donations (2 years)","(Nb donations (2 years))²", "Days since last donation", 
                        "iron supplementation", "Red meat","Vegetables", "Fruit and Berries", "Milk",
                         "Fruit Juices", "Coffee","Tea", "Beer", "Wine" , "Liquor"),          
          dep.var.caption   = "sTfR (log)",
          dep.var.labels.include = FALSE,
          column.labels = c("Pre-menopausal women", "Post-menopausal women", "Men"),
          # column.separate = c(2, 2,2),
          model.numbers= FALSE,
          report=('vcsp'),
          digits = 3)
```

-

### Supp Figure 3:  Visualizations of the effects of donation history on sTRF 

```{r, warning=FALSE}

plt_1 <-blood_data_summary_final %>%
  dplyr::select(donor, group2, TransferrinR, 
         TwoYearsFromStartCount_FB) %>% 
  ggplot(aes(x=TwoYearsFromStartCount_FB, y = TransferrinR)) +
  geom_hline(data = threshold, aes(yintercept = TransferrinR_thresh), color = "dark red", size = .2) +
  geom_jitter(aes(color = group2), alpha = 0.25, width = 0.2, size = 0.2, height = 0, shape = 1) +
  geom_boxplot(aes(color = group2, group = TwoYearsFromStartCount_FB), alpha = 0,
               width = 0.4,
               outlier.alpha = 0) +
  facet_grid(group2~.) +
  scale_color_manual(values = c( "#ff85a2",  "#de2d26","#00BFFF" ),
                     limits = c("Pre-menopausal women", "Post-menopausal women","Men" )) +
  theme_bw() +
  theme(legend.position = "none", 
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
       panel.border = element_rect(colour = NA),
        strip.background=element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(colour="black"),
        axis.title.x = element_text(size = 8),
        axis.title.y =element_text(size = 8),
        axis.text = element_text(size = 8)) +
  scale_y_log10() +
  scale_x_continuous(breaks=c(0, 2, 4,6,8, 10, 12)) +
  xlab("Nb of donations (2 years)") +
  ylab("sTfR (mg/l)") 

plt_2 <-blood_data_summary_final %>%
  dplyr::select(donor, group2, TransferrinR, 
         DaysToPreviousFB) %>% 
 ggplot(aes(x=DaysToPreviousFB, y=TransferrinR)) +
  geom_hline(data = threshold, aes(yintercept = TransferrinR_thresh), color = "dark red", size = .2) +
  geom_point(aes(color = group2), alpha = 0.4,  size = 0.7, shape =1) +
  facet_grid(group2~., 
      labeller = label_wrap_gen(width = 2, multi_line = TRUE))  +
  scale_color_manual(values = c( "#ff85a2",  "#de2d26","#00BFFF" ),
                     limits = c("Pre-menopausal women", "Post-menopausal women","Men" )) +
  theme_bw() +
 theme(legend.position = "none", 
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = NA),
        strip.background=element_rect(colour="white",fill="white"),
        strip.text = element_text(face="bold", size = 10),
        axis.line = element_line(colour="black"),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_blank(),
        axis.text = element_text(size = 8)) +
  scale_y_log10() +
  scale_x_log10(breaks=c(60, 90, 180, 360,1000)) +
  xlab("Days since last donation") +
  ylab("")


plt_3 <- blood_data_summary_final %>% 
  mutate(iron_comp_f = case_when(
    iron_comp_c == 3 ~ "All or almost all",
    iron_comp_c == 2 ~ "About half",
    iron_comp_c == 1 ~ "Less than half",
    iron_comp_c == 0 ~ "None",
    TRUE ~ "NA"),
  iron_comp_f = ordered(iron_comp_f, 
                  levels = c("NA","None", "Less than half",  "About half","All or almost all" ))) %>% 
  dplyr::select(group2, iron_comp_f, TransferrinR) %>% 
  ggplot(aes(x = iron_comp_f, y = TransferrinR, color = group2 )) +
  geom_hline(data = threshold, aes(yintercept = TransferrinR_thresh), color = "dark red", size = .2) +
  geom_boxplot(outlier.alpha = 0, position = "dodge", width = 0.5) +
  geom_jitter(width = 0.25, height = 0, shape = 1, size = 0.2, alpha = 0.25) +
  scale_color_manual(values = c( "#ff85a2",  "#de2d26","#00BFFF" ),
                     limits = c( "Pre-menopausal women", "Post-menopausal women", "Men" )) +
  facet_grid(group2~., 
      labeller = label_wrap_gen(width = 2, multi_line = TRUE)) +
 theme(legend.position = "none", 
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
       panel.border = element_rect(colour = NA),
        strip.background=element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(colour="black"),
        axis.text.x=element_text(angle=20,hjust=1),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        axis.text = element_text(size = 8)) +
  scale_y_log10(breaks=c(5, 15,25,50,100)) +
  xlab("Iron supplementation")  + 
  ylab("sTfR (mg/l)") 

plt_4 <- blood_data_summary_final %>% 
    mutate(red_meat = case_when(red_meat == "never" ~  "Never",
    red_meat == "less_than_once_weekly" ~ "Less than weekly",
    red_meat %in% c("1.3_week" , "4.6_week") ~ "Several weekly",
    TRUE   ~ "Daily or more" ), 
    red_meat = ordered(red_meat, 
                  levels = c("Never", "Less than weekly", "Several weekly", "Daily or more" ))) %>% 
  dplyr::select(group2, red_meat, TransferrinR) %>% 
  ggplot(aes(x = red_meat, y = TransferrinR, color = group2 )) +
  geom_hline(data = threshold, aes(yintercept = TransferrinR_thresh), color = "dark red", size = .2) +
  geom_boxplot(outlier.alpha = 0, position = "dodge", width = 0.5) +
  geom_jitter(width = 0.25, height = 0, shape = 1, size = 0.2, alpha = 0.3) +
  scale_color_manual(values = c( "#ff85a2",  "#de2d26","#00BFFF" ),
                     limits = c( "Pre-menopausal women", "Post-menopausal women", "Men" )) +
  facet_grid(group2~., 
      labeller = label_wrap_gen(width = 2, multi_line = TRUE)) +
 theme(legend.position = "none", 
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_rect(colour = NA),
        strip.background=element_rect(colour="white",fill="white"),
        strip.text = element_text(face="bold", size = 10),
        axis.line = element_line(colour="black"),
        axis.text.x=element_text(angle=20,hjust=1),
        axis.title.x =  element_text(size = 8),
       axis.title.y =  element_blank(),
        axis.text = element_text(size = 8)) +
  scale_y_log10() +
  xlab("Red meat") 

plt2 <- cowplot::plot_grid(plt_1, plt_2, plt_3, plt_4,  ncol = 2)

plt2

cowplot::save_plot("../results/figures/Supp_Figure_3_sTFR.png",plt2, base_height = NULL,
          base_aspect_ratio = .5, base_width = 11.5 * .394 ) # make room for figure legend
```


# Self-assessed health

Self-assessed health was measured using the following question:

* How is your health in general?
  * Poor, satisfactory, good, very good, excellent

We use ordinal logistic regression analysis [@Bender1997] to analyze the relationship between self-assessed health and ferritin. We test whether Ferritin level is a significant predictor of self-assessed health  when controlling for age, BMI, CRP, number of donations in the last two years, and physical activity. We run both a frequentist analysis and a Bayesian analysis. Results from the frequentist analysis suggest that Ferritin is not a significant predictor of self-perceived health in blood donors. According to the Bayesian analysis, the data dees not provide strong evidence in favor of ferritin co-varying with self-perceived health. 

## Supplementary Figure 4

We first display the distribution of answers to the self-reported health question.

```{r fig.show="80%"}
plt1 <- blood_data_summary_final %>% 
  dplyr::select(donor, group, self_perceived_health) %>% 
 mutate(self_perceived_health = ifelse(self_perceived_health == "Very_good", "Very good",
                                       as.character(self_perceived_health)),
         self_perceived_health = ordered(self_perceived_health, 
                                         levels = c("Satisfactory", "Good", "Very good", "Excellent"))) %>% 
  mutate(group = case_when(group == "Pre_menopause_women" ~ "Pre-menopausal \n women",
         group == "Post_menopause_women" ~ "Post-menopausal \n women",
         TRUE ~ "Men"),
         group = ordered(group, levels = c("Pre-menopausal \n women", "Post-menopausal \n women", "Men"))) %>% 
  group_by(group, self_perceived_health  ) %>% 
  summarise( N = n()) %>% 
  ungroup() %>% 
  group_by(group) %>% 
  mutate(proportion = round(N / sum(N), 2)) %>% 
  ggplot(aes(x = self_perceived_health, y = proportion, fill = group)) +
  # geom_jitter(alpha = 0.5, color = "light blue", shape = 1) +
  geom_col( color = "grey") + 
  geom_text(aes(label=round(proportion,2)), 
            position=position_dodge(width=0.9), 
            vjust=-0.5,
            size = 2) +
  facet_wrap( ~ group) +
  xlab("Question") +
  labs(y = "Proportion of donors") +
  scale_fill_manual(values = c( "#00BFFF",  "#de2d26", "#ff85a2" ),
                     limits = c( "Men",  "Post-menopausal \n women", "Pre-menopausal \n women" )) +
  scale_y_continuous(breaks = c(0, 0.1, 0.2, 0.3,  0.4, 0.5)) +
  xlab("How is your health in general?") +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 9, angle = 30, vjust = .75),
        strip.background = element_blank(),
        axis.title.x = element_text(size = 10),
        axis.title.y =element_text(size = 10),
        axis.text = element_text(size = 8),
        strip.text = element_text(size = 8)) 

plt1

ggplot2::ggsave("../results/figures/supp_figure_4_health.png", plt1,  height = 6.5 * 1.25, 
                width = 10 * 1.25, units = "cm", dpi = 300 )
```

## Figure 4 - Relationship between ferritin and self-reported health

```{r}

plt <- blood_data_summary_final %>%
  mutate(self_perceived_health = ifelse(self_perceived_health == "Very_good", "Very good",
                                        as.character(self_perceived_health)),
         self_perceived_health = ordered(self_perceived_health, 
                                         levels = c("Satisfactory", "Good", "Very good", "Excellent"))) %>% 
  mutate(group = case_when(group == "Pre_menopause_women" ~ "Pre-menopausal \n women",
         group == "Post_menopause_women" ~ "Post-menopausal \n women",
         TRUE ~ "Men"),
         group = ordered(group, levels = c("Pre-menopausal \n women", "Post-menopausal \n women", "Men"))) %>% 
  ggplot(aes(x = self_perceived_health, y = Ferritin, color = group)) +
  geom_hline(yintercept = 15, 
             color = "dark red", 
             size = .2,
             linetype = 2) +
  geom_boxplot(outlier.alpha = 0, 
               alpha = 0,
               size = 0.3,
               width = 0.5) +
  geom_jitter(alpha = 0.3, 
              shape = 1, 
              size = 0.15,
              position = position_jitterdodge(jitter.height = 0, 
                                              jitter.width = 0.5 )) +
   scale_color_manual(values = c( "#00BFFF",  "#de2d26", "#ff85a2" ),
                     limits = c( "Men",  "Post-menopausal \n women", "Pre-menopausal \n women" )) +
  facet_grid(group ~ .) +
  xlab("How is your health in general?") +
    ylab(expression("Ferritin ("*mu* "g/l)")) +
  scale_y_log10(breaks = c(5,15,25,50,100)) +
    theme_bw() +
  theme(legend.position = "none",
        legend.title = element_blank(),
        legend.background = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
       panel.border = element_rect(colour = NA),
        strip.background=element_blank(),
        axis.line = element_line(colour="black"),
        axis.title.x = element_text(size = 10),
        axis.title.y =element_text(size = 10),
        axis.text = element_text(size = 8)) 


plt

ggplot2::ggsave("../results/figures/figure_4_health.png", plt,  height = 8.5 * 1.25, 
                width = 6.38 * 1.25, units = "cm", dpi = 300 )

```

## Ordinal regression analysis

To analyze the answers to the self-assessed health question properly, we use an ordinal proportional odds model or ordinal logistic model. This type of model is used to model outcome variables that are ordered and represent an unknown latent variable. Here self-reported health is the ordered outcome variable we are trying to predict. Additional information about ordinal models in epidemiology can be found in [@abreu2009ordinal]. Because there are few donors who reported only satisfactory health, we pool together the data from pre-menopausal and post-menopausal women into a single women's group.

An example of how to interpret the odds-ratio that are computed from the regression coefficients can be found here [here](https://www.google.com/url?sa=t&rct=j&q=&esrc=s&source=web&cd=9&cad=rja&uact=8&ved=2ahUKEwjg1PO9g_TcAhVhD5oKHSt1BVYQFjAIegQIABAC&url=http%3A%2F%2Fweb.pdx.edu%2F~newsomj%2Fcdaclass%2Fho_ordinal%2520regression.pdf&usg=AOvVaw2NRuZfLiLnX4bUfqOPVUne). 

We transform Ferritin into *logFerritin = ln(Ferritin)/ln(2)*. With this transformation, an increase in one unit in logFerritin is equivalent to a doubling of Ferritin values. We also center and scale all variables. We also divide age by 5, so a unit increase in the age factor is actually an increase in 5 years in age. We center the data but do not scale it in order to make the coeffs easier to interpret.


### Ordinality of predictors

We must first check that a majority of our explanatory variables do in fact have an ordinal relationship to the outcome variable. 
It seems to be mosty the case.

```{r}
blood_data_summary_reg %>%
  dplyr::select(donor,  group, self_perceived_health, age, BMI, don_ct) %>% 
  gather(key = measure, value = value, -donor, -group, -self_perceived_health) %>% 
    mutate(self_perceived_health = ifelse(self_perceived_health == "Very_good", "Very good", 
                                          as.character(self_perceived_health)),
         self_perceived_health = ordered(self_perceived_health, 
                                         levels = c("Satisfactory", "Good", "Very good", "Excellent"))) %>% 
    mutate(group = case_when(group == "Pre_menopause_women" ~ "Pre-menopausal \n women",
         group == "Post_menopause_women" ~ "Post-menopausal \n women",
         TRUE ~ "Men"),
         measure = ifelse(measure == "don_ct", "Nb of \n donations", measure)) %>% 
  ggplot(aes(x = self_perceived_health, y = value, color = group)) +
  geom_boxplot(outlier.alpha = 0, alpha = 0) +
  geom_jitter(alpha = 0.2, 
              shape = 1, 
              size = 0.5,
              position = position_jitterdodge(jitter.height = 0, jitter.width = 0.5)) +
  facet_grid(measure ~ group, scales = "free_y") +
  theme(axis.text.x = element_text(angle = 30, vjust = 0.65)) +
  scale_color_manual(values = c( "#00BFFF",  "#de2d26", "#ff85a2" ),
                     limits = c( "Men",  "Post-menopausal \n women", "Pre-menopausal \n women" )) +
  xlab("How is your health in general?") +
   theme(legend.position = "none",
        legend.title = element_blank(),
        legend.background = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
       panel.border = element_rect(colour = NA),
        strip.background=element_blank(),
        axis.line = element_line(colour="black"),
        axis.title.x = element_text(size = 12),
        axis.title.y =element_blank(),
        axis.text = element_text(size = 8)) 

```


```{r}
blood_data_summary_reg %>%
  dplyr::select(donor,  group, self_perceived_health,  physical_act_daily_n, physical_act_freq_n) %>% 
  gather(key = measure, value = value, -donor, -group, -self_perceived_health) %>% 
  mutate(self_perceived_health = ifelse(self_perceived_health == "Very_good", "Very good", 
                                        as.character(self_perceived_health)),
         self_perceived_health = ordered(self_perceived_health, 
                                         levels = c("Satisfactory", "Good", "Very good", "Excellent"))) %>% 
  mutate(group = case_when(group == "Pre_menopause_women" ~ "Pre-menopausal \n women",
         group == "Post_menopause_women" ~ "Post-menopausal \n women",
         TRUE ~ "Men")) %>% 
  mutate(measure = ifelse(measure == "physical_act_daily_n", 
                          "Physical activity \n Daily amount", 
                          "Physical activity \n Frequency")) %>%
  ggplot(aes(x = self_perceived_health, y = value, color = group)) +
  geom_boxplot(outlier.alpha = 0, alpha = 0) +
  geom_jitter(alpha = 0.2, shape = 1, 
              size = 0.3,
              position = position_jitterdodge(jitter.height = 0.2, jitter.width = 0.5)) +
  facet_grid(measure ~ group, scales = "free_y") +
  theme(axis.text.x = element_text(angle = 30, vjust = 0.65)) +
  scale_color_manual(values = c( "#00BFFF",  "#de2d26", "#ff85a2" ),
                     limits = c( "Men",  "Post-menopausal \n women", "Pre-menopausal \n women" )) +
  xlab("How is your health in general?") +
   theme(legend.position = "none",
        legend.title = element_blank(),
        legend.background = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
       panel.border = element_rect(colour = NA),
        strip.background=element_blank(),
        axis.line = element_line(colour="black"),
        axis.title.x = element_text(size = 12),
        axis.title.y =element_blank(),
        axis.text = element_text(size = 8)) 
```

### Women 

```{r}

ordinal_data_temp <-
  blood_data_summary_reg %>% 
  filter(group != "Men") %>% 
  mutate(menopausal_status = ifelse(group == "Pre_menopause_women", 0, 1) ) %>% 
  mutate(age = age/5) %>% 
  dplyr::select(self_perceived_health, menopausal_status,  age,  BMI, log_CRP, don_ct,
                log_last_don,  Ferritin, TransferrinR, smoking, physical_act_daily_n,  physical_act_freq_n) %>% 
  rename(sTfR = TransferrinR) %>% 
  mutate(log_Ferritin = log(Ferritin)/log(2),
         age = scale(age, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         log_last_don = scale(log_last_don, scale = FALSE)[,1],
         don_ct = scale(don_ct, scale = FALSE)[,1],
         log_Ferritin = scale(log_Ferritin, scale = FALSE)[,1],
          sTfR = scale(sTfR, scale = FALSE)[,1],
         smoking = as.numeric(smoking) - 1)

```

#### Correlograms

```{r}

ggpairs(ordinal_data_temp, 
        columns = c("self_perceived_health", "age", "log_CRP",  "BMI", "don_ct", "log_Ferritin"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```

#### Frequentist fit

##### Assumption checks

The following plot is a visual tool to estimate if the proportional odds assumption is met. For each predictor variable, the assumption is met if the distance between the cross marker and the triangle marker is similar for all levels of the predictor variable. This method is detailed in  [@harrell2015ordinal]  as well as [here](ttps://stats.idre.ucla.edu/r/dae/ordinal-logistic-regression/). 


```{r}

X5_num <- as.integer(ordinal_data_temp$self_perceived_health)

sf <- function(y)
  c('Y>1' = qlogis(mean(y>1)),
    'Y>2' = qlogis(mean(y>2)),
    'Y>3' = qlogis(mean(y>3)))

s <- summary(X5_num ~ menopausal_status + age + BMI + log_CRP + don_ct  + log_last_don  +
               smoking +  physical_act_freq_n +  physical_act_daily_n + log_Ferritin,
             fun = sf, 
             data = ordinal_data_temp)

s[, 4] <- s[, 4] - s[, 3]
s[, 3] <- s[, 3] - s[, 3]

plot(s, which=1:3, pch=1:3, xlab='logit', main=' ', xlim=range(s[,3:4]))
```


##### Ordinal regression 

We fit the data using the rms R package. 

```{r}
ordinal_data_temp <-
 ordinal_data_temp %>%
  mutate(smoking = as.numeric(smoking)-1)

dd <- datadist(ordinal_data_temp) 
options(datadist="dd")


fit_women <- lrm(self_perceived_health ~ menopausal_status + age +  BMI + log_CRP + 
                  smoking +  don_ct + log_last_don  +
                  physical_act_daily_n + physical_act_freq_n  + log_Ferritin, 
           data = ordinal_data_temp,
              x = TRUE, y = TRUE)
```


```{r, results='asis'}
stargazer(fit_women,  
           title = "Women",
          intercept.bottom = FALSE,
          type = "html",
          single.row = TRUE,
          ci = TRUE,
          omit = "Constant",
          header = FALSE,
          report = "vcsp",
          covariate.labels = c("Health > Good","Health > Very good", " Health > Excellent", "Menopausal status",
                               "age", "BMI", "CRP (log)", "Smoking(yes)",
                               "Nb donations (2 years)", "Nb of days since last donation (log)", 
                                "Daily Physical Activity", "Physical activity frequency", "ln(Ferritin)/ln(2)"),
          dep.var.caption   = "Self-rated health", 
          dep.var.labels.include = FALSE,
          # column.labels = c("All observations", "Influential removed"),
          star.cutoffs = c(0.05, 0.01, 0.001))
```


Our model significantly predicts self-rated health (Pr(> chi2) <0.0001), but Ferritin is not a significant predictor (p= 0.078 ).

We now transform the regression coefficients into odds-ratios for easier interpretation of results.

```{r}
brm_coeffs <- tibble(Regressor = names(coef(fit_women)), Coefficient = coef(fit_women)) %>% 
  mutate(Odds.Ratio = exp(Coefficient))

brm_coeffs %>% 
   filter(!(Regressor %in% c("y>=Good", "y>=Very_good", "y>=Excellent"))) %>% 
  mutate(Regressor = dplyr::recode(Regressor, age = "Age",
                               log_CRP = "CRP",
                               smoking = "Smoking",
                               don_ct = "Nb of donations (2 years)",
                               log_last_don = "Days since last donation",
                               physical_act_daily_n = "Daily physical activity",
                               physical_act_freq_n = "Exercise frequency",
                               log_Ferritin = "Ferritin")) %>% 
    kable()
```

* An increase in 1 point of BMI age results in a `r round(brm_coeffs %>% filter(Regressor == "age") %>% magrittr::extract2("Odds.Ratio")-1,3)*-100`% decreases in your odds by of feeling in very good or excellent compared to satisfactory or good. 

* An increase in one scale point in exercise frequency results in an increase of `r round(brm_coeffs %>% filter(Regressor == "physical_act_freq_n") %>% magrittr::extract2("Odds.Ratio")-1,3)*100`% of your odds of felling in excellent health compared to very good, good or satisfactory health. 

* Once BMI, age, CRP, donation activity, smoking and physical activity are controlled for, there is a `r round(brm_coeffs %>% filter(Regressor == "log_Ferritin") %>% magrittr::extract2("Odds.Ratio")-1,3)*100`% increase in the odds of rating your health as good, very good or excellent compared to satisfactory for each 2-fold increase in Ferritin levels. This increase is not statistically significant.

##### Bootstrapped CIs

We use a bootstrap to estimate  95% BCa confidence intervals on the coefficients. 

```{r}
set.seed(5874)

output_file_distrib = ("../results/bootstraps/health_coeff_boot_object_std_seed_5874_women.rdata")
output_file_Bca = ("../results/bootstraps/health_coeff_boot_std_Bca_seed_5874_women.rdata")
# 

get_from_file = TRUE

if(get_from_file){
  load(output_file_distrib)
}else{
  seed_curr <- .Random.seed

  boot_object <- rms::bootcov(fit = fit_women, B = 10000, coef.reps = TRUE)

  save(boot_object, file = output_file_distrib)
}


if(get_from_file){
  load(output_file_Bca)
}else{
  conf_int_coeffs <- bootBCa(boot_object$boot.coef, 
        boot_object$boot.Coef, 
        type = 'bca', 
        n = boot_object$stats[[1]], 
        seed = seed_curr, 
        conf.int = 0.95)

  save(conf_int_coeffs, file = output_file_Bca)
}

```

```{r}
ordinal_regression_coefficients <- 
  as.data.frame(boot_object$coefficients) %>% 
  rownames_to_column() %>%
  filter(!rowname %in% c("y>=Good", "y>=Very_good", "y>=Excellent" )) 

names(ordinal_regression_coefficients) <- c("regressor", "coefficient")

conf_int_coeffs <- as.data.frame(conf_int_coeffs) 
names(conf_int_coeffs) <- names(boot_object$coefficients) 

ordinal_regression_coefficients_women <- 
  conf_int_coeffs %>% 
    rownames_to_column() %>% 
    gather(key = regressor, value = coefficient, -rowname) %>% 
    filter(!regressor %in% c("y>=Good", "y>=Very_good", "y>=Excellent" ) ) %>%
    rename(conf_int = rowname) %>% 
    spread(key = conf_int, value = coefficient) %>% 
    full_join(ordinal_regression_coefficients, by = "regressor") 

ordinal_regression_coefficients_women <-
  ordinal_regression_coefficients_women %>% 
    gather(key = key, value = value, -regressor) %>% 
    mutate(value = exp(value),
           key = paste(key, "OR", sep = "_")) %>% 
    spread(key = key, value =  value) %>% 
    full_join(ordinal_regression_coefficients_women, by = "regressor") %>% 
    mutate(sex = "Women",
         method = "GLM")
  
```


```{r}
## create list for stargazer table

CI_regression_health = list(2)
regress_coeff_health = list(2)

CI_regression_health[[1]] <- t(as.matrix(exp(conf_int_coeffs)))
regress_coeff_health[[1]] <- exp(boot_object$coefficients)

```

#### Ordinal model Bayesian fit

We use brms to run a bayesian fit of the same regression model. We compute LOO to show that adding ferritin does not improve the regression model. 


```{r}
ordinal_fit_B_women <- brm(formula = self_perceived_health ~  menopausal_status + age +  BMI + log_CRP + 
                                      smoking +  don_ct + log_last_don  +
                                      physical_act_daily_n + physical_act_freq_n  + log_Ferritin,  
                            data = ordinal_data_temp,
                            family = cumulative,
                            save_all_pars = TRUE)
```




```{r message=FALSE, warning=FALSE}
bayes_regression_coeffs_women <- 
  tidy(ordinal_fit_B_women) %>% 
  dplyr::select(-std.error) %>% 
   dplyr::filter(!term %in% c("b_Intercept[1]", "b_Intercept[2]", "b_Intercept[3]","lp__",
                            "temp_Intercept[1]", "temp_Intercept[2]", "temp_Intercept[3]" )) %>%
  mutate(term = str_remove(term, "b_")) %>% 
  rename(regressor = term, coefficient = estimate, 
         Lower = lower, Upper = upper)

bayes_regression_coeffs_women <-
   bayes_regression_coeffs_women %>% 
    gather(key = key, value = value, -regressor) %>% 
    mutate(value = exp(value),
           key = paste(key, "OR", sep = "_")) %>% 
    spread(key = key, value =  value) %>% 
    full_join(bayes_regression_coeffs_women, by = "regressor") %>% 
   mutate(sex = "Women",
         method = "Bayes") %>%    
    mutate(regressor = ifelse(regressor == "smokingyes", "smoking_yes", regressor )) 

```


##### Model diagnostics

###### Posterior probability check

```{r message=FALSE, warning=FALSE}
pp_check(ordinal_fit_B_women)
```

###### Diagnostics of the MCMC sampling

```{r message=FALSE, warning=FALSE, fig.show="80%"}

stan_trace(ordinal_fit_B_women$fit)
# raftery.diag(data.brms)
stan_ac(ordinal_fit_B_women$fit)
stan_rhat(ordinal_fit_B_women$fit)
stan_ess(ordinal_fit_B_women$fit)
```


##### Model comparison 

Model comparison using Bayes Factors in non trivial when using ordinal models. We use LOO (Leave-One-Out cross validation) as is shown [here](https://kevinstadler.github.io/blog/bayesian-ordinal-regression-with-random-effects-using-brms/). [@vehtari2017practical] is the main reference. 
We compare a restricted model without Ferritin as a predictor and a full model with Ferritin as a predictor. 

```{r message=FALSE, warning=FALSE, }
ordinal_fit_restricted <- brm(formula = self_perceived_health ~ menopausal_status + age + BMI +  
                                        log_CRP + smoking + don_ct   +
                                        log_last_don + physical_act_daily_n + physical_act_freq_n ,
                              data = ordinal_data_temp,
                              family = cumulative,
                              save_all_pars = TRUE)

```

```{r}
LOO(ordinal_fit_restricted, ordinal_fit_B_women) 

```

The LOO comparison suggests that the full model does not fit the data any better than the restricted one since the LOO difference is of smaller magnitude than its SE. 


### Men

```{r}

ordinal_data_temp <-
  blood_data_summary_reg %>% 
  filter(group == "Men") %>% 
  mutate(age = age/5) %>% 
  dplyr::select(self_perceived_health,  age,  BMI, log_CRP, don_ct, log_last_don,  Ferritin, smoking, 
                physical_act_daily_n,  physical_act_freq_n) %>% 
  mutate(log_Ferritin = log(Ferritin)/log(2),
         age = scale(age, scale = FALSE)[,1],
         BMI = scale(BMI, scale = FALSE)[,1],
         log_CRP = scale(log_CRP, scale = FALSE)[,1],
         log_last_don = scale(log_last_don, scale = FALSE)[,1],
         don_ct = scale(don_ct, scale = FALSE)[,1],
         log_Ferritin = scale(log_Ferritin, scale = FALSE)[,1],
         smoking = as.numeric(smoking) - 1)

```


#### Correlograms


```{r}

ggpairs(ordinal_data_temp, 
        columns = c("self_perceived_health", "age", "log_CRP",  "BMI", "don_ct", "log_Ferritin"),
         lower = list(continuous = wrap("points", alpha = 0.3,size=0.1),
                      combo = wrap("facethist", binwidth = 0.5)),
        progress = FALSE)
```

#### Frequentist fit

##### Assumption checks


```{r}
X5_num <- as.integer(ordinal_data_temp$self_perceived_health)

sf <- function(y)
  c('Y>1' = qlogis(mean(y>1)),
    'Y>2' = qlogis(mean(y>2)),
    'Y>3' = qlogis(mean(y>3)))

s <- summary(X5_num ~ age + BMI + log_CRP + don_ct + log_last_don + 
               log_Ferritin + smoking +  
               physical_act_freq_n +  physical_act_daily_n,
             fun = sf, 
             data = ordinal_data_temp)

s[, 4] <- s[, 4] - s[, 3]
s[, 3] <- s[, 3] - s[, 3]
s[, 2] <- s[, 3] - s[, 2]


plot(s, which=1:3, pch=1:3, xlab='logit', main=' ', xlim=range(s[,2:4]))
```


##### Ordinal regression fit 

```{r}
ordinal_data_temp <-
 ordinal_data_temp %>%
  mutate(smoking = as.numeric(smoking)-1)

dd <- datadist(ordinal_data_temp) 
options(datadist="dd")
options(contrasts=c("contr.treatment", "contr.treatment"))

fit_men <- lrm(self_perceived_health ~  age +  BMI + log_CRP + smoking +  don_ct + log_last_don + 
             physical_act_daily_n + physical_act_freq_n  + log_Ferritin  , 
           data = ordinal_data_temp,
              x = TRUE, y = TRUE)
```


```{r, results='asis'}
stargazer(fit_men,  
           title = "Men",
          intercept.bottom = FALSE,
          type = "html",
          single.row = TRUE,
          ci = TRUE,
          omit = "Constant",
          header = FALSE,
          report = "vcsp",
          covariate.labels = c("Health > Good","Health > Very Good", "Health > Excellent",
                               "age", "BMI", "CRP (log)", "Smoking(yes)",
                               "Nb donations (2 years)", "Nb of days since last donation (log)", 
                               "Daily Physical Activity", "Physical activity frequency", "ln(Ferritin)/ln(2)"),
           dep.var.caption   = "Self-rated health", 
          dep.var.labels.include = FALSE,
          # column.labels = c("All observations", "Influential removed"),
          star.cutoffs = c(0.05, 0.01, 0.001))
```


Our model significantly predicts self-rated health (Pr(> chi2) <0.0001), but Ferritin is not a significant predictor (p= 0.2115 ).

We can now transform the regression coefficients into odds-ratios for easier interpretation of results.

```{r}
brm_coeffs <- tibble(Regressor = names(coef(fit_men)), Coefficient = coef(fit_men)) %>% 
  mutate(Odds.Ratio = exp(Coefficient))

brm_coeffs %>% 
   filter(!(Regressor %in% c("y>=Good", "y>=Very_good", "y>=Excellent"))) %>% 
  mutate(Regressor = dplyr::recode(Regressor, age = "Age",
                               log_CRP = "CRP",
                               smoking = "Smoking",
                               don_ct = "Nb of donations (2 years)",
                               log_last_don = "Days since last donation",
                               physical_act_daily_n = "Daily physical activity",
                               physical_act_freq_n = "Exercise frequency",
                               log_Ferritin = "Ferritin")) %>% 
    kable()
```

* An increase in 1 point of BMI age results in a `r round(brm_coeffs %>% filter(Regressor == "BMI") %>% magrittr::extract2("Odds.Ratio")-1,3)*-100` % decrease in your odds by of feeling in very good pr excellent compared to satisfactory or good. 

* An increase in one scale point in exercise frequency results in an increase of `r round(brm_coeffs %>% filter(Regressor == "physical_act_freq_n") %>% magrittr::extract2("Odds.Ratio")-1,3)*100`% of your odds of felling in excellent health compared to very good, good or satisfactory health. 

* Once BMI, age, CRP, donation activity, smoking and physical activity are controlled for, there is a `r round(brm_coeffs %>% filter(Regressor == "log_Ferritin") %>% magrittr::extract2("Odds.Ratio")-1,3)*100`% increase in the odds of rating your health as good, very good or excellent compared to satisfactory for each 2-fold increase in Ferritin levels. This increase is not statistically significant.

##### Bootstrapped CIs

We use a bootstrap to estimate  confidence intervals on the coefficients for proper reporting of results. 

```{r}
set.seed(5874)

output_file_distrib = ("../results/bootstraps/health_coeff_boot_object_std_seed_5874_men.rdata")
output_file_Bca = ("../results/bootstraps/health_coeff_boot_std_Bca_seed_5874_men.rdata")

get_from_file = TRUE

if(get_from_file){
  load(output_file_distrib)
}else{
  seed_curr <- .Random.seed

  boot_object <- rms::bootcov(fit = fit_men, B = 10000, coef.reps = TRUE)

  save(boot_object, file = output_file_distrib)
}


if(get_from_file){
  load(output_file_Bca)
}else{
  conf_int_coeffs <- bootBCa(boot_object$boot.coef, 
        boot_object$boot.Coef, 
        type = 'bca', 
        n = boot_object$stats[[1]], 
        seed = seed_curr, 
        conf.int = 0.95)

  save(conf_int_coeffs, file = output_file_Bca)
}

```

```{r}
ordinal_regression_coefficients <- 
  as.data.frame(boot_object$coefficients) %>% 
  rownames_to_column() %>%
  filter(!rowname %in% c("y>=Good", "y>=Very_good", "y>=Excellent" )) 

names(ordinal_regression_coefficients) <- c("regressor", "coefficient")

conf_int_coeffs <- as.data.frame(conf_int_coeffs) 
names(conf_int_coeffs) <- names(boot_object$coefficients) 

ordinal_regression_coefficients_men <- 
  conf_int_coeffs %>% 
    rownames_to_column() %>% 
    gather(key = regressor, value = coefficient, -rowname) %>% 
    filter(!regressor %in% c("y>=Good", "y>=Very_good", "y>=Excellent" ) ) %>% 
    rename(conf_int = rowname) %>% 
    spread(key = conf_int, value = coefficient) %>% 
    full_join(ordinal_regression_coefficients, by = "regressor") 

ordinal_regression_coefficients_men <-
  ordinal_regression_coefficients_men %>% 
    gather(key = key, value = value, -regressor) %>% 
    mutate(value = exp(value),
           key = paste(key, "OR", sep = "_")) %>% 
    spread(key = key, value =  value) %>% 
    full_join(ordinal_regression_coefficients_men, by = "regressor") %>% 
    mutate(sex = "Men",
         method = "GLM")
  
```

```{r}
## create list for stargazer table
CI_regression_health[[2]] <- t(as.matrix(exp(conf_int_coeffs)))
regress_coeff_health[[2]] <- exp(boot_object$coefficients)

```

#### Ordinal model Bayesian fit

```{r}
ordinal_fit_B_men <- brm(formula = self_perceived_health ~  age +  BMI + log_CRP + smoking + 
                                  don_ct + log_last_don  +
                                  physical_act_daily_n + physical_act_freq_n  + log_Ferritin,  
                         data = ordinal_data_temp,
                         family = cumulative,
                         save_all_pars = TRUE)

```



```{r message=FALSE, warning=FALSE}
bayes_regression_coeffs_men <- 
  tidy(ordinal_fit_B_men) %>% 
  dplyr::select(-std.error) %>% 
   dplyr::filter(!term %in% c("b_Intercept[1]", "b_Intercept[2]", "b_Intercept[3]","lp__",
                            "temp_Intercept[1]", "temp_Intercept[2]", "temp_Intercept[3]" )) %>%
  mutate(term = str_remove(term, "b_")) %>% 
  rename(regressor = term, coefficient = estimate, 
         Lower = lower, Upper = upper)

bayes_regression_coeffs_men <-
   bayes_regression_coeffs_men %>% 
    gather(key = key, value = value, -regressor) %>% 
    mutate(value = exp(value),
           key = paste(key, "OR", sep = "_")) %>% 
    spread(key = key, value =  value) %>% 
    full_join(bayes_regression_coeffs_men, by = "regressor") %>% 
   mutate(sex = "Men",
         method = "Bayes") %>%    
    mutate(regressor = ifelse(regressor == "smokingyes", "smoking_yes", regressor )) 

```

##### Model diagnostics

###### Posterior probability check

```{r message=FALSE, warning=FALSE}
pp_check(ordinal_fit_B_men)
```

###### Diagnostics of the MCMC sampling

```{r message=FALSE, warning=FALSE, fig.show="80%"}

stan_trace(ordinal_fit_B_men$fit)
# raftery.diag(data.brms)
stan_ac(ordinal_fit_B_men$fit)
stan_rhat(ordinal_fit_B_men$fit)
stan_ess(ordinal_fit_B_men$fit)
```


##### Model comparison 

We use LOO (Leave-One-Out cross validation). We compare a restricted model without Ferritin as a predictor and a full model with Ferritin as a predictor. 

```{r message=FALSE, warning=FALSE}
ordinal_fit_restricted <- brm(formula = self_perceived_health ~ age + BMI +  
                                        log_CRP + smoking + don_ct   +
                                        log_last_don + physical_act_daily_n + physical_act_freq_n ,
                              data = ordinal_data_temp,
                              family = cumulative,
                              save_all_pars = TRUE)

```

```{r message=FALSE, warning=FALSE}
LOO(ordinal_fit_restricted, ordinal_fit_B_men)
```

Adding Ferritin does not improve the fit between the data and the model. 


## Supplementary Tables 

### Ordinal logistic regression 

```{r , results='asis'}

stargazer(fit_women, fit_men,
          out = "../results/tables/supp_table_3_frequentist_health.html",
          intercept.bottom = FALSE,
          title = "Ordinal logistic regression",
          type = "html",
          single.row = TRUE,
          ci = TRUE,
          ci.custom = CI_regression_health,
          coef = regress_coeff_health,
          p.auto = FALSE,
          header = FALSE,
          covariate.labels =c("Health > Good", "Health > y VeGood", "Health > Excellent", "Menopausal status",
                          "Age",  "BMI", "CRP", "Smoking(yes)", "Nb donations (2 years)",
                          "Nb of days since last donation",
                        "Physical activity (daily amount)", "Physical activity (frequency)", "Ferritin"),
          dep.var.caption   = "How has your health been in general?", 
          dep.var.labels.include = FALSE,
          column.labels = c("Women", "Men"),
          # column.separate = c(2, 2,2),
          model.numbers= FALSE,
          report=('vcsp'))
```


### Bayesian ordinal logistic regression


```{r, results='asis'}
bayes_regression_coeffs_women %>% 
  bind_rows(bayes_regression_coeffs_men) %>% 
  rename(group = sex) %>% 
  dplyr::select(-coefficient:-Upper, - method) %>% 
  mutate(CI = paste(round(Lower_OR,2), round(Upper_OR,2), sep = "-" )) %>% 
  dplyr::select(group, regressor, coefficient_OR, CI ) %>% 
  mutate(regressor = dplyr::recode(regressor, age = "Age",
                       don_ct = "Nb donations (2 years)",
                       log_CRP = "CRP",
                       log_Ferritin = "Ferritin",
                       log_last_don =  "Nb of days since last donation",
                       menopausal_status = "Menopausal status",
                       physical_act_daily_n = "Physical activity (daily amount)", 
                       physical_act_freq_n = "Physical activity (frequency)",
                       smoking = "Smoking")) %>% 
  rename(Odds.Ratio = coefficient_OR) %>% 
  stargazer(title= "Bayesian ordinal logistic regression",
            type = "html", 
          summary = FALSE, 
          rownames = FALSE,
          header = TRUE,
          out = "../results/tables/supp_table_4_ordinal_bayes_health.html",
          initial.zero = FALSE, 
          digits = 2)  
```


### Summary table


```{r}
final_results <- ordinal_regression_coefficients_women %>% 
  bind_rows(bayes_regression_coeffs_women, 
            ordinal_regression_coefficients_men,
            bayes_regression_coeffs_men) %>% 
  filter(regressor == "log_Ferritin") %>% 
  mutate(regressor = "Ferritin") %>% 
  rename(Odds.Ratio = coefficient_OR,
         Odds.Ratio.low.CI = Lower_OR,
         Odds.Ratio.upp.CI = Upper_OR,
         coeff.low.CI = Lower,
          coeff.upp.CI = Upper) %>% 
  dplyr::select(regressor:Odds.Ratio.upp.CI, coefficient,coeff.low.CI,coeff.upp.CI )
  
```

```{r , results='asis'}
stargazer(final_results, 
          type = "html", 
          summary = FALSE, 
          rownames = FALSE,
          header = TRUE,
          out = "../results/tables/table_4_ordinal_reg.html",
          initial.zero = FALSE, 
          digits = 5)
```


# References 